---
title: "Compare CASAL & Casal2 for the HOK assessment"
author: "C.Marsh"
date: "15 November 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r wrap-hook, include=FALSE}
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})
```
## Comparing CASAL mpd with Casal2 models
Comparing CASAL and Casal2 models for the modified stock assessment. This model differs from the actual assessment by two key points. First sex is ignored in this assessment, and secondly the original assessment uses length based selectivities.


The firsrt thing we want to confirm is that when run deterministically with the exact same parameters (\texttt{casal2 -r -i casal_pars.txt > run.log}) that they give the same initial partition, SSB and expected values as CASAL. You should always look to confirm if the initial age or length structure is correct. If this isn't correct it suggests that either growth, movement or natural mortality are incorrectly specified (assuming you are using an equilibrium age structure). 


```{r source, linewidth=60}
library(casal)
library(casal2)
library(xtable)
## bring in functions
source("AuxillaryFunsToincludeInLibrary.r")

#cas_mpd = casal::extract.mpd("../HOK/CASAL/estimate.log")
cas_mpd = casal::extract.mpd("../HOK/CASAL/run.log")
#cas_orig = casal::extract.mpd("../HOK/CASAL/orig_run.log")
cas2_mpd = casal2::extract.mpd("../HOK/Casal2/run.log")

## look at B0
cas_mpd$quantities$B0
cas2_mpd$Recruit_E$`1`$b0
cas2_mpd$Recruit_W$`1`$b0

## look at R0
cas_mpd$quantities$R0$E - cas2_mpd$Recruit_E$`1`$r0
cas_mpd$quantities$R0$W - cas2_mpd$Recruit_W$`1`$r0

# look at m
cas_mpd$free$natural_mortality.all
cas2_mpd$Mortality$`1`$m

# plot SSB
cas2_ssb = plot.derived_quantities(cas2_mpd, "SSB", plot.it = F)
par(mfrow = c(1,2))
plot(rownames(cas2_ssb), cas2_ssb[,"SSB_E"], ylim = c(0,1e6), lwd = 2, type = "l",
     xlab = "years", ylab = "SSB", main = "Eastern stock")
lines(rownames(cas2_ssb), cas_mpd$quantities$SSBs$E, lty = 2, lwd = 2, col = "red")


plot(rownames(cas2_ssb), cas2_ssb[,"SSB_W"], ylim = c(0,1.6e6), lwd = 2, type = "l",
     xlab = "years", ylab = "SSB", main = "Western stock")
lines(rownames(cas2_ssb), cas_mpd$quantities$SSBs$W, lty = 2, lwd = 2, col = "red")
legend('bottomleft', legend = c("Casal2","CASAL"),col = c("black","red"), lwd = 2, cex = 0.6)

## Recruitment
par(mfrow = c(1,2))
plot(cas_mpd$quantities$true_YCS$year, cas_mpd$quantities$true_YCS$E,
     type = "l", lwd = 2, col = "red", ylab  = "True YCS", xlab = "years", 
     ylim = c(0,3), main = "Eastern")
lines(cas2_mpd$Recruit_E$`1`$ycs_years, cas2_mpd$Recruit_E$`1`$true_ycs, 
      col = "black", lwd =2 ,lty = 2)
legend('topright', legend = c("Casal2","CASAL"),col = c("black","red"), lwd = 2, cex = 0.6)

plot(cas_mpd$quantities$true_YCS$year, cas_mpd$quantities$true_YCS$W, 
     type = "l", lwd = 2, col = "red", ylab  = "", xlab = "years", 
     ylim = c(0,3), main = "Western")
lines(cas2_mpd$Recruit_W$`1`$ycs_years, cas2_mpd$Recruit_W$`1`$true_ycs, col = "black", lwd =2,
      lty = 2)

```

So this looks like the process dynamics are doing the right thing, lets look at how Casal2 is generating fits to observations, and the likelihood contribution.

```{r obs, linewidth=60, results="asis"}
cas_obj = cas_mpd$objective.function$components
cas2_obj = split_obj(cas2_mpd, label = "objective") ## custom function,
sum(cas2_obj$Value) - cas2_mpd$objective$`1`$values["total_score"] 
cas2_mpd$objective$`1`$values["total_score"] 
obs_across_both = sum(cas_obj$label %in% cas2_obj$Label)
comp = matrix(NA, nrow = obs_across_both, ncol = 3)
colnames(comp) = c("CASAL", "Casal2","Diff")
labels = cas_obj$label[cas_obj$label %in% cas2_obj$Label]
nonlabels = cas_obj$label[!cas_obj$label %in% cas2_obj$Label]

rownames(comp) = labels
for(i in 1:obs_across_both) {
  comp[i,1] = cas_obj$value[cas_obj$label == labels[i]]
  comp[i,2] = cas2_obj$Value[cas2_obj$Label == labels[i]]
  comp[i,3] = comp[i,1] - comp[i,2]
}
tab = xtable(comp, digits = 8, caption = "Comparison of Casal2 and CASAL
             objective function contributions for the main data sets.")
print(tab)
```


# Compare Casal2 and CASAL estimation
```{r est, linewidth=60, results="asis"}
cas_mpd = casal::extract.mpd("../HOK/CASAL/estimate.log")
cas2_mpd = casal2::extract.mpd("../HOK/Casal2/estimate.log")

## look at B0
cas_mpd$quantities$B0
cas2_mpd$Recruit_E$`1`$b0
cas2_mpd$Recruit_W$`1`$b0

## look at R0
cas_mpd$quantities$R0$E - cas2_mpd$Recruit_E$`1`$r0
cas_mpd$quantities$R0$W - cas2_mpd$Recruit_W$`1`$r0

# look at m
cas_mpd$free$natural_mortality.all
cas2_mpd$Mortality$`1`$m

# plot SSB
cas2_ssb = plot.derived_quantities(cas2_mpd, "SSB", plot.it = F)
par(mfrow = c(1,2))
plot(rownames(cas2_ssb), cas2_ssb[,"SSB_E"], ylim = c(0,1e6), lwd = 2, type = "l",
     xlab = "years", ylab = "SSB", main = "Eastern stock")
lines(rownames(cas2_ssb), cas_mpd$quantities$SSBs$E, lty = 2, lwd = 2, col = "red")


plot(rownames(cas2_ssb), cas2_ssb[,"SSB_W"], ylim = c(0,1.6e6), lwd = 2, type = "l",
     xlab = "years", ylab = "SSB", main = "Western stock")
lines(rownames(cas2_ssb), cas_mpd$quantities$SSBs$W, lty = 2, lwd = 2, col = "red")
legend('bottomleft', legend = c("Casal2","CASAL"),col = c("black","red"), lwd = 2, cex = 0.6)

## Recruitment
par(mfrow = c(1,2))
plot(cas_mpd$quantities$true_YCS$year, cas_mpd$quantities$true_YCS$E,
     type = "l", lwd = 2, col = "red", ylab  = "True YCS", xlab = "years", 
     ylim = c(0,3), main = "Eastern")
lines(cas2_mpd$Recruit_E$`1`$ycs_years, cas2_mpd$Recruit_E$`1`$true_ycs, 
      col = "black", lwd =2 ,lty = 2)
legend('topright', legend = c("Casal2","CASAL"),col = c("black","red"), lwd = 2, cex = 0.6)

plot(cas_mpd$quantities$true_YCS$year, cas_mpd$quantities$true_YCS$W, 
     type = "l", lwd = 2, col = "red", ylab  = "", xlab = "years", 
     ylim = c(0,3), main = "Western")
lines(cas2_mpd$Recruit_W$`1`$ycs_years, cas2_mpd$Recruit_W$`1`$true_ycs, col = "black", lwd =2,
      lty = 2)

```

So this looks like the process dynamics are doing the right thing, lets look at how Casal2 is generating fits to observations, and the likelihood contribution.

```{r obs_est, linewidth=60, results="asis"}
cas_obj = cas_mpd$objective.function$components
cas2_obj = split_obj(cas2_mpd, label = "objective") ## custom function,
sum(cas2_obj$Value) - cas2_mpd$objective$`1`$values["total_score"] 
cas2_mpd$objective$`1`$values["total_score"] 
obs_across_both = sum(cas_obj$label %in% cas2_obj$Label)
comp = matrix(NA, nrow = obs_across_both, ncol = 3)
colnames(comp) = c("CASAL", "Casal2","Diff")
labels = cas_obj$label[cas_obj$label %in% cas2_obj$Label]
nonlabels = cas_obj$label[!cas_obj$label %in% cas2_obj$Label]

rownames(comp) = labels
for(i in 1:obs_across_both) {
  comp[i,1] = cas_obj$value[cas_obj$label == labels[i]]
  comp[i,2] = cas2_obj$Value[cas2_obj$Label == labels[i]]
  comp[i,3] = comp[i,1] - comp[i,2]
}
tab = xtable(comp, digits = 8, caption = "Comparison of Casal2 and CASAL
             objective function contributions for the main data sets.")
print(tab)
```




