---
title: "Casal2 Test Case MPD comparisons for SCI"
output:
    html_document:
        toc: TRUE
        toc_depth: 2
        number_sections: TRUE

header-includes:
 - \usepackage{pdflscape}
 - \newcommand{\blandscape}{\begin{landscape}}
 - \newcommand{\elandscape}{\end{landscape}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# SCI comparison of CASAL and Casal2 model configurations

This document compares the results of at least 2 CASAL model configurations (base and at least one sensitivity) and up to 6 Casal2 model configurations (3 BetaDiff and 3 ADOL-C).

The CASAL model sensitivity 1 has a smaller minimisation tolerance value than the CASAL base model (1e-8 vs. 2e-6).

The Casal2 ADOL-C and BetaDiff low tolerance models have the same tolerance values than the CASAL base model (1e-8 vs. 2e-6).


## SCI model characteristics

The main characteristics of the Test Case SCI (Scampi) CASAL model are:

* length-based sex-specific model
* years 1986 - 2022, projection years 2023 - 2027
* three time steps, Oct - Jan, Feb - Apr, May - Sep
* estimate growth transition matrix
* four tag releases
* Beverton-Holt stock-recruitment relationship, with steepness (h) 0.75 and $CV_R$ 0.7 (prior)
* length-weight relationship ($W = aL^b$)
* three fisheries

Observation data include:

* trawl fishery CPUE
* trawl survey abundance and proportions-at-length
* fishery proportions-at-length
* tag-recapture data for growth estimation
* maturity data

Parameters estimated include:

* B0
* sex-specific growth parameters
* maturity parameters
* YCS (recruitment deviations)
* sex-specific fishery and trawl selectivity parameters


## R environment

```{r C1_C2_setup, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_set_up_R_env.R')
```


## CASAL and Casal2 model output

```{r CASAL, warning=FALSE}
source('../../R-functions/report_read_in_CASAL_MPD_files.R')
```


```{r Casal2}
source('../../R-functions/report_read_in_Casal2_MPD_files.R')
```



\newpage
\blandscape

## Tables

Tables of parameter estimates and objective function components for the CASAL and Casal2 model MPD results

```{r table_setup, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_set_up_tables.R')

options(scipen=999)
```


```{r huxtables_parameters, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_print_parameter_tables.R')

hux_C1_est_params
hux_C2_est_params
hux_C2_pd_est_params
```


```{r huxtables_obj_fun, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_print_obj_fun_tables.R')

hux_C1_obj_fun
hux_C2_obj_fun
```


```{r convergence_info, echo=FALSE}
source('../../R-functions/report_print_convergence_info.R')
```


```{r warnings_and_bounds, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_print_warnings.R')
```


```{r huxtables_close, echo=FALSE}
options(scipen=0)
```


\elandscape
\newpage



## Matching of outputs

Time series comparisons with CASAL base model results

```{r Time_Series_Match, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_time_series_match_function.R')


c1_catch <- c1_quant$actual_catches$Trawl_1

for (c in 1:num_C2_models)
{
    print(paste('Catch time series base model comparison for run', C2_subdir[c]))

    print(paste('Actual catches for Trawl 1 match:', time_series_match(c1_catch, cas2_mpd[[c]]$Instantaneous_Mortality$`catch[Trawl_1]`)))

    print('')
}


c1_catch <- c1_quant$actual_catches$Trawl_2

for (c in 1:num_C2_models)
{
    print(paste('Catch time series base model comparison for run', C2_subdir[c]))

    print(paste('Actual catches for Trawl 2 match:', time_series_match(c1_catch, cas2_mpd[[c]]$Instantaneous_Mortality$`catch[Trawl_2]`)))

    print('')
}


c1_catch <- c1_quant$actual_catches$Trawl_3

for (c in 1:num_C2_models)
{
    print(paste('Catch time series base model comparison for run', C2_subdir[c]))

    print(paste('Actual catches for Trawl 3 match:', time_series_match(c1_catch, cas2_mpd[[c]]$Instantaneous_Mortality$`catch[Trawl_3]`)))

    print('')
}
```

Derived quantities

SB0, SBcurrent, MSY, F_MSY, others...



## Plots

Comparison plots

```{r plots_base, echo=FALSE}

# c(bottom, left, top, right)
par(mar=c(4,4,2,1) + 0.1)

par(mfrow=c(1,2))

bl_lwd <- 6
gr_lwd <- 3
```


```{r plots_correlation, echo=FALSE}
source('../../R-functions/report_plot_correlation_matrices.R')
```


```{r plots_set_0, echo=FALSE}
source('../../R-functions/report_set_up_legend.R')


max_val <- max(c1_quant$SSBs$SSB,
               c1_sens1_quant$SSBs$SSB,
               max(unlist(list.map(cas2_mpd, max(SSB$SSB$values)))))

plot(c1_quant$SSBs$year, c1_quant$SSBs$SSB, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='Biomass (t)', main='SSB comparison')
lines(c1_sens1_quant$SSBs$year, c1_sens1_quant$SSBs$SSB, type='l', col='grey', lwd=gr_lwd)

for (c in 1:num_C2_models)
{
    lines(names(cas2_mpd[[c]]$SSB$SSB$values), cas2_mpd[[c]]$SSB$SSB$values, col=C2_color[c], lwd=1)
}

plot_legend()
```



```{r plots_set_1, echo=FALSE}

# Q:  how to get CASAL initial age structure?
# A:  use 'casal::extract.partition()''


lengths <- as.numeric(colnames(cas2_mpd[[1]]$Init$values))

# omit the 'category' column
# 'category' column is no longer output by Casal2 as of 2022-05-16

# C1 cas_part$initial[1,] is for untagged males, cas_part$initial[2,] is for untagged females
# C2 cas2_mpd[[N]]$Init$values[1,] is for untagged males, cas2_mpd[[N]]$Init$values[5,] is for untagged females


c1_rows <- c(1, 2)
c2_rows <- c(1, 5)

c2_mat <- array(0, dim=c(num_C2_models, dim(as.matrix(cas2_mpd[[1]]$Init$values[c2_rows,]))))
for (c in 1:num_C2_models)
{
    c2_mat[c,,] <- as.matrix(cas2_mpd[[c]]$Init$values[c2_rows,])
}

# plot initial numbers

max_val <- max(cas_part$initial,
               cas_sens1_part$initial,
               max(c2_mat))

plot(lengths, cas_part$initial[1,], type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Length', ylab='', main='Initial numbers-at-length comparison (untagged)', sub='Dotted lines for female')
lines(lengths, cas_part$initial[2,], col='black', lwd=bl_lwd, lty=3)

lines(lengths, cas_sens1_part$initial[1,], col='grey', lwd=gr_lwd)
lines(lengths, cas_sens1_part$initial[2,], col='grey', lwd=gr_lwd, lty=3)

for (c in 1:num_C2_models)
{
    lines(lengths, c2_mat[c,1,], col=C2_color[c], lwd=1)
    lines(lengths, c2_mat[c,2,], col=C2_color[c], lwd=1, lty=3)
}

plot_legend()
```


```{r plots_set_2, echo=FALSE}

c2_mat <- array(0, dim=c(num_C2_models, dim(as.matrix(cas2_mpd[[1]]$state_Oct_Jan$`1986`$values[c2_rows,]))))
for (c in 1:num_C2_models)
{
    c2_mat[c,,] <- as.matrix(cas2_mpd[[c]]$state_Oct_Jan$`1986`$values[c2_rows,])
}

# plot initial year numbers (time step 1)

max_val <- max(cas_part$initial,
               cas_sens1_part$initial,
               max(c2_mat))

plot(lengths, cas_part$initial[1,], type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Length', ylab='', main='1986 numbers-at-length comparison (untagged, time step 1, Oct-Jan)', sub='Dotted lines for female')
lines(lengths, cas_part$initial[2,], col='black', lwd=bl_lwd, lty=3)

lines(lengths, cas_sens1_part$initial[1,], col='grey', lwd=gr_lwd)
lines(lengths, cas_sens1_part$initial[2,], col='grey', lwd=gr_lwd, lty=3)

for (c in 1:num_C2_models)
{
    lines(lengths, c2_mat[c,1,], col=C2_color[c], lwd=1)
    lines(lengths, c2_mat[c,2,], col=C2_color[c], lwd=1, lty=3)
}

plot_legend()


c2_mat <- array(0, dim=c(num_C2_models, dim(as.matrix(cas2_mpd[[1]]$state_May_Sep$`2022`$values[c2_rows,]))))
for (c in 1:num_C2_models)
{
    c2_mat[c,,] <- as.matrix(cas2_mpd[[c]]$state_May_Sep$`2022`$values[c2_rows,])
}

# plot final year numbers (time step 3)

max_val <- max(cas_part$final,
               cas_sens1_part$final,
               max(c2_mat))

plot(lengths, cas_part$final[1,], type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Length', ylab='', main='2022 numbers-at-length comparison (untagged, time step 3, May-Sep)', sub='Dotted lines for female')
lines(lengths, cas_part$final[2,], col='black', lwd=bl_lwd, lty=3)

lines(lengths, cas_sens1_part$final[1,], col='grey', lwd=gr_lwd)
lines(lengths, cas_sens1_part$final[2,], col='grey', lwd=gr_lwd, lty=3)

for (c in 1:num_C2_models)
{
    lines(lengths, c2_mat[c,1,], col=C2_color[c], lwd=1)
    lines(lengths, c2_mat[c,2,], col=C2_color[c], lwd=1, lty=3)
}

plot_legend()
```


```{r plots_set_3, echo=FALSE}
# plot YCS

max_val <- max(c1_quant$`Vector parameter values`$recruitment.YCS,
               c1_sens1_quant$`Vector parameter values`$recruitment.YCS,
               max(unlist(list.map(cas2_mpd, max(Recruitment$ycs_values)))))

plot(c1_quant$`Vector parameter values`$recruitment.YCS, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='', main='YCS comparison')
lines(c1_sens1_quant$`Vector parameter values`$recruitment.YCS, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Recruitment$ycs_values, col=C2_color[c], lwd=1)
}

plot_legend()



# plot true YCS

max_val <- max(c1_quant$true_YCS$true_YCS,
               c1_sens1_quant$true_YCS$true_YCS,
               max(unlist(list.map(cas2_mpd, max(Recruitment$true_ycs)))))

plot(c1_quant$true_YCS$year, c1_quant$true_YCS$true_YCS, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='', main='true YCS comparison')
lines(c1_sens1_quant$true_YCS$year, c1_sens1_quant$true_YCS$true_YCS, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Recruitment$ycs_years, cas2_mpd[[c]]$Recruitment$true_ycs, col=C2_color[c], lwd=1)
}

plot_legend()



# plot recruitment

max_val <- max(c1_quant$recruitments$recruitment,
               c1_sens1_quant$recruitments$recruitment,
               max(unlist(list.map(cas2_mpd, max(Recruitment$Recruits)))))

plot(c1_quant$recruitments$year, c1_quant$recruitments$recruitment, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='', main='Recruitment comparison')
lines(c1_sens1_quant$recruitments$year, c1_sens1_quant$recruitments$recruitment, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines((cas2_mpd[[c]]$Recruitment$ycs_years + cas2_mpd[[c]]$Recruitment$ssb_offset), cas2_mpd[[c]]$Recruitment$Recruits, col=C2_color[c], lwd=1)
}

plot_legend()
```


```{r plots_set_4, echo=FALSE}
# plot fishery CPUE
# c1_quant$`CPUE-Trawl_1.fits`
# cas2_mpd[[1]]$$`CPUE-Trawl_1`

c1_surv    <- unlist(c1_quant$`CPUE-Trawl_1.fits`)
c1_s1_surv <- unlist(c1_sens1_quant$`CPUE-Trawl_1.fits`)

c2_surv <- list()
for (c in 1:num_C2_models)
{
    c2_surv[[c]] <- cas2_mpd[[c]]$`CPUE-Trawl_1`$Values
}

max_val <- max(c1_surv,
               c1_s1_surv,
               (c2_surv[[1]]$observed * exp(1.96 * c2_surv[[1]]$error_value)),
               max(unlist(list.map(c2_surv, max(expected)))))

plot(names(c1_surv), c1_surv, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='Biomass (t)', main='Trawl fishery 1 CPUE index comparison', sub='Compare process error estimates')
lines(names(c1_s1_surv), c1_s1_surv, type='l', col='grey', lwd=gr_lwd)
points(c2_surv[[1]]$year, c2_surv[[1]]$observed, pch=20, col='black')
arrows(c2_surv[[1]]$year, (c2_surv[[1]]$observed * exp(-1.96 * c2_surv[[1]]$error_value)), c2_surv[[1]]$year, (c2_surv[[1]]$observed * exp(1.96 * c2_surv[[1]]$error_value)), length=0.05, angle=90, code=3)
for (c in 1:num_C2_models)
{
    points(c2_surv[[c]]$year, c2_surv[[c]]$expected, col=C2_color[c], pch=(15+c-1))
}

plot_legend()
```


```{r plots_set_5, echo=FALSE}
# plot surveys
# c1_quant$TrawlSurvey_1.fits and c1_quant$TrawlSurvey_1.fits
# cas2_mpd[[1]]$TrawlSurvey_1 and cas2_mpd[[1]]$TrawlSurvey_1

c1_surv    <- unlist(c1_quant$TrawlSurvey_1.fits)
c1_s1_surv <- unlist(c1_sens1_quant$TrawlSurvey_1.fits)

c2_surv <- list()
for (c in 1:num_C2_models)
{
    c2_surv[[c]] <- cas2_mpd[[c]]$TrawlSurvey_1$Values
}

max_val <- max(c1_surv,
               c1_s1_surv,
               (c2_surv[[1]]$observed * exp(1.96 * c2_surv[[1]]$error_value)),
               max(unlist(list.map(c2_surv, max(expected)))))

plot(names(c1_surv), c1_surv, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='Biomass (t)', main='Trawl survey 1 index comparison')
lines(names(c1_s1_surv), c1_s1_surv, type='l', col='grey', lwd=gr_lwd)
points(c2_surv[[1]]$year, c2_surv[[1]]$observed, pch=20, col='black')
arrows(c2_surv[[1]]$year, (c2_surv[[1]]$observed * exp(-1.96 * c2_surv[[1]]$error_value)), c2_surv[[1]]$year, (c2_surv[[1]]$observed * exp(1.96 * c2_surv[[1]]$error_value)), length=0.05, angle=90, code=3)
for (c in 1:num_C2_models)
{
    points(c2_surv[[c]]$year, c2_surv[[c]]$expected, col=C2_color[c], pch=(15+c-1))
}

plot_legend()


c1_surv    <- unlist(c1_quant$TrawlSurvey_2.fits)
c1_s1_surv <- unlist(c1_sens1_quant$TrawlSurvey_2.fits)

c2_surv <- list()
for (c in 1:num_C2_models)
{
    c2_surv[[c]] <- cas2_mpd[[c]]$TrawlSurvey_2$Values
}

max_val <- max(c1_surv,
               c1_s1_surv,
               (c2_surv[[1]]$observed * exp(1.96 * c2_surv[[1]]$error_value)),
               max(unlist(list.map(c2_surv, max(expected)))))

plot(names(c1_surv), c1_surv, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='Biomass (t)', main='Trawl survey 2 index comparison')
lines(names(c1_s1_surv), c1_s1_surv, type='l', col='grey', lwd=gr_lwd)
points(c2_surv[[1]]$year, c2_surv[[1]]$observed, pch=20, col='black')
arrows(c2_surv[[1]]$year, (c2_surv[[1]]$observed * exp(-1.96 * c2_surv[[1]]$error_value)), c2_surv[[1]]$year, (c2_surv[[1]]$observed * exp(1.96 * c2_surv[[1]]$error_value)), length=0.05, angle=90, code=3)
for (c in 1:num_C2_models)
{
    points(c2_surv[[c]]$year, c2_surv[[c]]$expected, col=C2_color[c], pch=(15+c-1))
}

plot_legend()
```


```{r plots_set_6, echo=FALSE}
# > names(cas2_mpd[[1]])
#  [1] "Init"                            "state_Oct_Jan"                   "state_Feb_Apr"                   "state_May_Sep"                   "estimated_values"
#  [6] "estimated_summary"               "obj_fun"                         "Hess"                            "Covar"                           "Corr"
# [11] "Recruitment"                     "Instantaneous_Mortality"         "maturity"                        "FishingSel_1_male"               "FishingSel_1_female"
# [16] "FishingSel_2_male"               "FishingSel_2_female"             "TrawlSurveySel_1_male"           "TrawlSurveySel_1_female"         "TrawlSurveySel_2_male"
# [21] "TrawlSurveySel_2_female"         "CPUE-Trawl_1"                    "TrawlSurvey_1"                   "TrawlSurvey_2"                   "__CPUE-Commercialq_1__"
# [26] "__TrawlSurveyq_1__"              "__TrawlSurveyq_2__"              "SSB"                             "TrawlSurveyProportionAtLength_1" "TrawlSurveyProportionAtLength_2"
# [31] "CommercialCatchLength_1"         "CommercialCatchLength_2"         "CommercialCatchLength_3"         "Cryer_Oliver_maturity"           "1997_recaptures1_1"
# [36] "1997_recaptures1_2"              "1997_recaptures2_1"              "1997_recaptures2_2"              "1997_recaptures2_3"              "1997_recaptures3_1"
# [41] "1997_recaptures3_2"              "1997_recaptures3_3"              "growth"                          "tag_1996_2_male"                 "tag_1996_2_female"
# [46] "tag_1996_3_male"                 "tag_1996_3_female"               "tag_1997_1_male"                 "tag_1997_1_female"               "One"
# [51] "tag_1996_2_male.one"             "tag_1996_2_female.one"           "tag_1996_3_male.one"             "tag_1996_3_female.one"           "tag_1997_1_male.one"
# [56] "tag_1997_1_female.one"           "messages_encountered"
# > names(cas2_mpd[[1]]$Instantaneous_Mortality)
#  [1] "biomass"                    "categories"
#  [3] "label"                      "m"
#  [5] "relative_m_by_length"       "time_step_proportions"
#  [7] "sub_type"                   "year"
#  [9] "fishing_pressure[Trawl_1]"  "exploitation_rate[Trawl_1]"
# [11] "catch[Trawl_1]"             "actual_catch[Trawl_1]"
# [13] "fishing_pressure[Trawl_2]"  "exploitation_rate[Trawl_2]"
# [15] "catch[Trawl_2]"             "actual_catch[Trawl_2]"
# [17] "fishing_pressure[Trawl_3]"  "exploitation_rate[Trawl_3]"
# [19] "catch[Trawl_3]"             "actual_catch[Trawl_3]"
# [21] "type"

# plot fishing pressures

max_val <- max(c1_quant$fishing_pressures$Trawl_1,
               c1_sens1_quant$fishing_pressures$Trawl_1,
               max(unlist(list.map(cas2_mpd, max(Instantaneous_Mortality$`fishing_pressure[Trawl_1]`)))))

plot(c1_quant$fishing_pressures$year, c1_quant$fishing_pressures$Trawl_1, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='U', main='Fishing pressure comparison: Trawl 1')
lines(c1_sens1_quant$fishing_pressures$year, c1_sens1_quant$fishing_pressures$Trawl_1, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Instantaneous_Mortality$year, cas2_mpd[[c]]$Instantaneous_Mortality$`fishing_pressure[Trawl_1]`, col=C2_color[c], lwd=1)
}

plot_legend()


max_val <- max(c1_quant$fishing_pressures$Trawl_2,
               c1_sens1_quant$fishing_pressures$Trawl_2,
               max(unlist(list.map(cas2_mpd, max(Instantaneous_Mortality$`fishing_pressure[Trawl_2]`)))))

plot(c1_quant$fishing_pressures$year, c1_quant$fishing_pressures$Trawl_2, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='U', main='Fishing pressure comparison: Trawl 2')
lines(c1_sens1_quant$fishing_pressures$year, c1_sens1_quant$fishing_pressures$Trawl_2, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Instantaneous_Mortality$year, cas2_mpd[[c]]$Instantaneous_Mortality$`fishing_pressure[Trawl_2]`, col=C2_color[c], lwd=1)
}

plot_legend()


max_val <- max(c1_quant$fishing_pressures$Trawl_3,
               c1_sens1_quant$fishing_pressures$Trawl_3,
               max(unlist(list.map(cas2_mpd, max(Instantaneous_Mortality$`fishing_pressure[Trawl_3]`)))))

plot(c1_quant$fishing_pressures$year, c1_quant$fishing_pressures$Trawl_3, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='U', main='Fishing pressure comparison: Trawl 3')
lines(c1_sens1_quant$fishing_pressures$year, c1_sens1_quant$fishing_pressures$Trawl_3, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Instantaneous_Mortality$year, cas2_mpd[[c]]$Instantaneous_Mortality$`fishing_pressure[Trawl_3]`, col=C2_color[c], lwd=1)
}

plot_legend()
```


```{r plots_set_7, echo=FALSE}
# > names(c1_quant$`Ogive parameter values`)
# [1] "maturity_props.all"
# [2] "selectivity[FishingSel_1].female"
# [3] "selectivity[FishingSel_2].female"
# [4] "selectivity[TrawlSurveySel_1].female"
# [5] "selectivity[TrawlSurveySel_2].female"
# [6] "selectivity[FishingSel_1].male"
# [7] "selectivity[FishingSel_2].male"
# [8] "selectivity[TrawlSurveySel_1].male"
# [9] "selectivity[TrawlSurveySel_2].male"

# cas2_mpd[[1]]$maturity$Value

max_val <- max(1,
               c1_quant$`Ogive parameter values`$maturity_props.all,
               c1_sens1_quant$`Ogive parameter values`$maturity_props.all)

plot(lengths, c1_quant$`Ogive parameter values`$maturity_props.all, type='l', col='black', lwd=bl_lwd, ylim=c(0,max_val), xlab='Length', ylab='', main='Maturity-at-length comparison')
lines(lengths, c1_sens1_quant$`Ogive parameter values`$maturity_props.all, col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(lengths, cas2_mpd[[c]]$maturity$Values, col=C2_color[c], lwd=1)
}

plot_legend()
```


```{r plots_set_8, echo=FALSE}

max_val <- max(1,
               c1_quant$`Ogive parameter values`$`selectivity[FishingSel_1].male`,
               c1_quant$`Ogive parameter values`$`selectivity[FishingSel_1].female`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[FishingSel_1].male`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[FishingSel_1].female`)

plot(lengths, c1_quant$`Ogive parameter values`$`selectivity[FishingSel_1].male`, type='l', col='black', lwd=bl_lwd, ylim=c(0,max_val), xlab='Length', ylab='', main='Trawl fishery 1 selectivity-at-length comparison', , sub='Dotted lines for female')
lines(lengths, c1_quant$`Ogive parameter values`$`selectivity[FishingSel_1].female`, type='l', col='black', lwd=bl_lwd, lty=3)

lines(lengths, c1_sens1_quant$`Ogive parameter values`$`selectivity[FishingSel_1].male`, col='grey', lwd=gr_lwd)
lines(lengths, c1_sens1_quant$`Ogive parameter values`$`selectivity[FishingSel_1].female`, col='grey', lwd=gr_lwd, lty=3)

for (c in 1:num_C2_models)
{
    lines(lengths, cas2_mpd[[c]]$FishingSel_1_male$Values, col=C2_color[c], lwd=1)
    lines(lengths, cas2_mpd[[c]]$FishingSel_1_female$Values, col=C2_color[c], lwd=1, lty=3)
}

plot_legend()


max_val <- max(1,
               c1_quant$`Ogive parameter values`$`selectivity[FishingSel_2].male`,
               c1_quant$`Ogive parameter values`$`selectivity[FishingSel_2].female`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[FishingSel_2].male`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[FishingSel_2].female`)

plot(lengths, c1_quant$`Ogive parameter values`$`selectivity[FishingSel_2].male`, type='l', col='black', lwd=bl_lwd, ylim=c(0,max_val), xlab='Length', ylab='', main='Trawl fishery 2 selectivity-at-length comparison', , sub='Dotted lines for female')
lines(lengths, c1_quant$`Ogive parameter values`$`selectivity[FishingSel_2].female`, type='l', col='black', lwd=bl_lwd, lty=3)

lines(lengths, c1_sens1_quant$`Ogive parameter values`$`selectivity[FishingSel_2].male`, col='grey', lwd=gr_lwd)
lines(lengths, c1_sens1_quant$`Ogive parameter values`$`selectivity[FishingSel_2].female`, col='grey', lwd=gr_lwd, lty=3)

for (c in 1:num_C2_models)
{
    lines(lengths, cas2_mpd[[c]]$FishingSel_2_male$Values, col=C2_color[c], lwd=1)
    lines(lengths, cas2_mpd[[c]]$FishingSel_2_female$Values, col=C2_color[c], lwd=1, lty=3)
}

plot_legend()
```


```{r plots_set_9, echo=FALSE}

max_val <- max(1,
               c1_quant$`Ogive parameter values`$`selectivity[TrawlSurveySel_1].male`,
               c1_quant$`Ogive parameter values`$`selectivity[TrawlSurveySel_1].female`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[TrawlSurveySel_1].male`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[TrawlSurveySel_1].female`)

plot(lengths, c1_quant$`Ogive parameter values`$`selectivity[TrawlSurveySel_1].male`, type='l', col='black', lwd=bl_lwd, ylim=c(0,max_val), xlab='Length', ylab='', main='Trawl survey 1 selectivity-at-length comparison', , sub='Dotted lines for female')
lines(lengths, c1_quant$`Ogive parameter values`$`selectivity[TrawlSurveySel_1].female`, type='l', col='black', lwd=bl_lwd, lty=3)

lines(lengths, c1_sens1_quant$`Ogive parameter values`$`selectivity[TrawlSurveySel_1].male`, col='grey', lwd=gr_lwd)
lines(lengths, c1_sens1_quant$`Ogive parameter values`$`selectivity[TrawlSurveySel_1].female`, col='grey', lwd=gr_lwd, lty=3)

for (c in 1:num_C2_models)
{
    lines(lengths, cas2_mpd[[c]]$TrawlSurveySel_1_male$Values, col=C2_color[c], lwd=1)
    lines(lengths, cas2_mpd[[c]]$TrawlSurveySel_1_female$Values, col=C2_color[c], lwd=1, lty=3)
}

plot_legend()


max_val <- max(1,
               c1_quant$`Ogive parameter values`$`selectivity[TrawlSurveySel_2].male`,
               c1_quant$`Ogive parameter values`$`selectivity[TrawlSurveySel_2].female`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[TrawlSurveySel_2].male`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[TrawlSurveySel_2].female`)

plot(lengths, c1_quant$`Ogive parameter values`$`selectivity[TrawlSurveySel_2].male`, type='l', col='black', lwd=bl_lwd, ylim=c(0,max_val), xlab='Length', ylab='', main='Trawl survey 2 selectivity-at-length comparison', , sub='Dotted lines for female')
lines(lengths, c1_quant$`Ogive parameter values`$`selectivity[TrawlSurveySel_2].female`, type='l', col='black', lwd=bl_lwd, lty=3)

lines(lengths, c1_sens1_quant$`Ogive parameter values`$`selectivity[TrawlSurveySel_2].male`, col='grey', lwd=gr_lwd)
lines(lengths, c1_sens1_quant$`Ogive parameter values`$`selectivity[TrawlSurveySel_2].female`, col='grey', lwd=gr_lwd, lty=3)

for (c in 1:num_C2_models)
{
    lines(lengths, cas2_mpd[[c]]$TrawlSurveySel_2_male$Values, col=C2_color[c], lwd=1)
    lines(lengths, cas2_mpd[[c]]$TrawlSurveySel_2_female$Values, col=C2_color[c], lwd=1, lty=3)
}

plot_legend()
```


```{r plots_set_10, echo=FALSE}
# code from Samik Datta and Vidette McGregor
source('../../R-functions/growth_curve_fn.R')

nyears <- 40

# values from CASAL population.csl and Casal2 Population.csl2 files
l_alpha <- 20
l_beta  <- 40
cv_val  <- 0.1 # 0.0 in CASAL file, 0.1 in Casal2 file - which one to use?
initial_size_mean <- 10.0
initial_size_cv   <- 0.40


# C1
g_vec     <- c1_quant$`Vector parameter values`$`growth[1].g_male`
min_sigma <- c1_quant$`Scalar parameter values`$`growth[1].minsigma_male`

pars_tib <- tibble('g1' = g_vec[1], 'g2' = g_vec[2], 'l1' = l_alpha, 'l2' = l_beta, 'cv' = cv_val, 'min_sigma' = min_sigma)
size_obj <- growth_curve_fn(pars = pars_tib, length_vec = lengths, nyears = nyears, sex = 'male', initial_length = initial_size_mean, initial_cv = initial_size_cv)

all_size_obj <- bind_cols(size_obj, 'model' = 'CASAL', 'lwd' = bl_lwd)


g_vec     <- c1_quant$`Vector parameter values`$`growth[1].g_female`
min_sigma <- c1_quant$`Scalar parameter values`$`growth[1].minsigma_male`

pars_tib <- tibble('g1' = g_vec[1], 'g2' = g_vec[2], 'l1' = l_alpha, 'l2' = l_beta, 'cv' = cv_val, 'min_sigma' = min_sigma)
size_obj <- growth_curve_fn(pars = pars_tib, length_vec = lengths, nyears = nyears, sex = 'female', initial_length = initial_size_mean, initial_cv = initial_size_cv)

all_size_obj <- bind_rows(all_size_obj, bind_cols(size_obj, 'model' = 'CASAL', 'lwd' = bl_lwd))


# C1 sensitivity
g_vec     <- c1_sens1_quant$`Vector parameter values`$`growth[1].g_male`
min_sigma <- c1_sens1_quant$`Scalar parameter values`$`growth[1].minsigma_male`

pars_tib <- tibble('g1' = g_vec[1], 'g2' = g_vec[2], 'l1' = l_alpha, 'l2' = l_beta, 'cv' = cv_val, 'min_sigma' = min_sigma)
size_obj <- growth_curve_fn(pars = pars_tib, length_vec = lengths, nyears = nyears, sex = 'male', initial_length = initial_size_mean, initial_cv = initial_size_cv)

all_size_obj <- bind_rows(all_size_obj, bind_cols(size_obj, 'model' = 'CASAL sens1', 'lwd' = gr_lwd))


g_vec     <- c1_sens1_quant$`Vector parameter values`$`growth[1].g_female`
min_sigma <- c1_sens1_quant$`Scalar parameter values`$`growth[1].minsigma_male`

pars_tib <- tibble('g1' = g_vec[1], 'g2' = g_vec[2], 'l1' = l_alpha, 'l2' = l_beta, 'cv' = cv_val, 'min_sigma' = min_sigma)
size_obj <- growth_curve_fn(pars = pars_tib, length_vec = lengths, nyears = nyears, sex = 'female', initial_length = initial_size_mean, initial_cv = initial_size_cv)

all_size_obj <- bind_rows(all_size_obj, bind_cols(size_obj, 'model' = 'CASAL sens1', 'lwd' = gr_lwd))


# C2
for (c in 1:num_C2_models)
{
    g_alpha   <- cas2_mpd[[c]]$estimated_values$values$`growth_increment[male_growth].g_alpha`
    g_beta    <- cas2_mpd[[c]]$estimated_values$values$`growth_increment[male_growth].g_beta`
    min_sigma <- cas2_mpd[[c]]$estimated_values$values$`growth_increment[female_growth].min_sigma`

    pars_tib <- tibble('g1' = g_alpha, 'g2' = g_beta, 'l1' = l_alpha, 'l2' = l_beta, 'cv' = cv_val, 'min_sigma' = min_sigma)
    size_obj <- growth_curve_fn(pars = pars_tib, length_vec = lengths, nyears = nyears, sex = 'male', initial_length = initial_size_mean, initial_cv = initial_size_cv)

    all_size_obj <- bind_rows(all_size_obj, bind_cols(size_obj, 'model' = paste('Casal2', C2_subdir[c]), 'lwd' = 1))


    g_alpha   <- cas2_mpd[[c]]$estimated_values$values$`growth_increment[female_growth].g_alpha`
    g_beta    <- cas2_mpd[[c]]$estimated_values$values$`growth_increment[female_growth].g_beta`
    min_sigma <- cas2_mpd[[c]]$estimated_values$values$`growth_increment[female_growth].min_sigma`

    pars_tib <- tibble('g1' = g_alpha, 'g2' = g_beta, 'l1' = l_alpha, 'l2' = l_beta, 'cv' = cv_val, 'min_sigma' = min_sigma)
    size_obj <- growth_curve_fn(pars = pars_tib, length_vec = lengths, nyears = nyears, sex = 'female', initial_length = initial_size_mean, initial_cv = initial_size_cv)

    all_size_obj <- bind_rows(all_size_obj, bind_cols(size_obj, 'model' = paste('Casal2', C2_subdir[c]), 'lwd' = 1))
}



# plot
all_color_vec <- c('black', 'grey', C2_color)

p <- all_size_obj %>%
        mutate(Age = as.integer(Time)) %>%
        ggplot(aes(x = Age, y = mean_length, colour=model)) +
        geom_path(stat = "identity", aes(size=lwd)) + facet_grid(sex ~ .) +
        guides(size = "none") + labs(title="Mean length at age by sex") +
        scale_colour_brewer(palette = "Set1")

p


# plot_legend()
```
