---
title: "Casal2 Test Case comparisons for LIN"
output:
    html_document:
        toc: TRUE
        toc_depth: 2
        number_sections: TRUE

header-includes:
 - \usepackage{pdflscape}
 - \newcommand{\blandscape}{\begin{landscape}}
 - \newcommand{\elandscape}{\end{landscape}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# LIN comparison of CASAL and Casal2 model configurations

This document compares the results of at least 2 CASAL model configurations (base and at least one sensitivity) and up to 8 Casal2 model configurations (3 BetaDiff, 2 CppAD, and 3 ADOL-C).

The CASAL model sensitivity 1 has a smaller tolerance value than the CASAL base model (1e-6 vs. 2e-3).

The Casal2 ADOL-C and BetaDiff low tolerance models have a smaller tolerance value than the CASAL base model (1e-6 vs. 2e-3). The Casal2 CppAD models have a tolerance value of 1e-9.


## LIN model characteristics

The main characteristics of the Test Case LIN (ling) CASAL model are:

* one stock, ages 3 - 25
* one area
* years 1972 - 2018, projection years 2019 - 2023
* two time steps: Dec_Aug and Sep_Nov
* two stock categories: male and female
* natural mortality-at-age is double exponential
* sex-specific maturity ogives specified by age
* sex-specific von Bertalanffy age-length relationship
* sex-specific length-weight relationship ($W = aL^b$)
* Beverton-Holt stock-recruitment relationship, with steepness (h) 0.84
* ageing error
* two surveys with sex-specific logistic selectivity
* three fisheries, one with double normal selectivity and two with logistic selectivity

Observation data include:

* survey indices
* survey proportions-at-age
* fishery proportions-at-age

Parameters estimated include:

* B0
* parameters for the natural mortality-at-age ogive
* catchability (q) for the two surveys
* parameters for the sex-specific selectivity curves for the two surveys
* parameters for the sex-specific selectivity curves for the three fisheries
* YCS (recruitment deviations)


## R environment

```{r C1_C2_setup, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_set_up_R_env.R')
```


## CASAL and Casal2 model output

```{r CASAL, warning=FALSE}
source('../../R-functions/report_read_in_CASAL_MPD_files.R')
```


```{r Casal2}
source('../../R-functions/report_read_in_Casal2_MPD_files.R')
```



\newpage
\blandscape

## Tables

Tables of parameter estimates and objective function components for the CASAL and Casal2 model MPD results

```{r table_setup, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_set_up_tables.R')

options(scipen=999)
```


```{r huxtables_parameters, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_print_parameter_tables.R')

hux_C1_est_params
hux_C2_est_params
hux_C2_pd_est_params
```


```{r huxtables_obj_fun, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_print_obj_fun_tables.R')

hux_C1_obj_fun
hux_C2_obj_fun
```


```{r convergence_info, echo=FALSE}
source('../../R-functions/report_print_convergence_info.R')
```


```{r warnings_and_bounds, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_print_warnings.R')
```


```{r huxtables_close, echo=FALSE}
options(scipen=0)
```

\elandscape
\newpage



## Matching of outputs

Time series comparisons with CASAL base model results

```{r Time_Series_Match, echo=FALSE, warning=FALSE, message=FALSE}
# do the catch time series differ?

source('../../R-functions/report_time_series_match_function.R')


for (c in 1:num_C2_models)
{
    print(paste('Catch time series base model comparison for run', C2_subdir[c]))

    print(paste('Actual catches for trawl match:',
                time_series_match(c1_quant$actual_catches$trawl, cas2_mpd[[c]]$Mortality$`catch[FishingTrwl]`)))

    print(paste('Actual catches for line_home match:',
                time_series_match(c1_quant$actual_catches$line_home, cas2_mpd[[c]]$Mortality$`catch[FishingLineHome]`)))

    print(paste('Actual catches for line_spawn match:',
                time_series_match(c1_quant$actual_catches$line_spawn, cas2_mpd[[c]]$Mortality$`catch[FishingLineSpawn]`)))

    print('')
}
```

Derived quantities

SB0, SBcurrent, MSY, F_MSY, others...



## Plots

Comparison plots

```{r plots_base, echo=FALSE}

# c(bottom, left, top, right)
par(mar=c(4,4,2,1) + 0.1)

par(mfrow=c(1,2))

bl_lwd <- 6
gr_lwd <- 3
```


```{r plots_correlation, echo=FALSE}
source('../../R-functions/report_plot_correlation_matrices.R')
```


```{r plots_set_0, echo=FALSE}
source('../../R-functions/report_set_up_legend.R')


max_val <- max(c1_quant$SSBs$SSB,
               c1_sens1_quant$SSBs$SSB,
               max(unlist(list.map(cas2_mpd, max(DerivedQuantities$SSB$values)))))

plot(c1_quant$SSBs$year, c1_quant$SSBs$SSB, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='Biomass (t)', main='SSB comparison')
lines(c1_sens1_quant$SSBs$year, c1_sens1_quant$SSBs$SSB, type='l', col='grey', lwd=gr_lwd)

for (c in 1:num_C2_models)
{
    lines(names(cas2_mpd[[c]]$DerivedQuantities$SSB$values), cas2_mpd[[c]]$DerivedQuantities$SSB$values, col=C2_color[c], lwd=1)
}

plot_legend()
```


```{r plots_set_1, echo=FALSE}

# Q:  how to get CASAL initial age structure?
# A:  use 'casal::extract.partition()''


ages <- seq(3, 25)

# omit the 'category' column
c2_mat <- array(0, dim=c(num_C2_models, dim(as.matrix(cas2_mpd[[1]]$Init$values[,-1]))))
for (c in 1:num_C2_models)
{
    c2_mat[c,,] <- as.matrix(cas2_mpd[[c]]$Init$values[,-1])
}

# plot initial numbers

max_val <- max(cas_part$initial,
               cas_sens1_part$initial,
               max(c2_mat))

plot(ages, cas_part$initial[1,], type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Age', ylab='', main='Initial numbers-at-age comparison (male and female)')
lines(ages, cas_part$initial[2,], col='black', lwd=bl_lwd, lty=3)

lines(ages, cas_sens1_part$initial[1,], col='grey', lwd=gr_lwd)
lines(ages, cas_sens1_part$initial[2,], col='grey', lwd=gr_lwd, lty=3)

for (c in 1:num_C2_models)
{
    lines(ages, c2_mat[c,1,], col=C2_color[c], lwd=1)
    lines(ages, c2_mat[c,2,], col=C2_color[c], lwd=1, lty=3)
}

plot_legend()



c2_mat <- array(0, dim=c(num_C2_models, dim(as.matrix(cas2_mpd[[1]]$state1$`2018`$values))))
for (c in 1:num_C2_models)
{
    c2_mat[c,,] <- as.matrix(cas2_mpd[[c]]$state1$`2018`$values)
}

# plot final year numbers

max_val <- max(cas_part$final,
               cas_sens1_part$final,
               max(c2_mat))

plot(ages, cas_part$final[1,], type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Age', ylab='', main='Final numbers-at-age comparison (male and female) (time step 1)')
lines(ages, cas_part$final[2,], col='black', lwd=bl_lwd, lty=3)

lines(ages, cas_sens1_part$final[1,], col='grey', lwd=gr_lwd)
lines(ages, cas_sens1_part$final[2,], col='grey', lwd=gr_lwd, lty=3)

for (c in 1:num_C2_models)
{
    lines(ages, c2_mat[c,1,], col=C2_color[c], lwd=1)
    lines(ages, c2_mat[c,2,], col=C2_color[c], lwd=1, lty=3)
}

plot_legend()



c2_mat <- array(0, dim=c(num_C2_models, dim(as.matrix(cas2_mpd[[1]]$state2$`2018`$values))))
for (c in 1:num_C2_models)
{
    c2_mat[c,,] <- as.matrix(cas2_mpd[[c]]$state2$`2018`$values)
}

# plot final year numbers

max_val <- max(cas_part$final,
               cas_sens1_part$final,
               max(c2_mat))

plot(ages, cas_part$final[1,], type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Age', ylab='', main='Final numbers-at-age comparison (male and female) (time step 2)')
lines(ages, cas_part$final[2,], col='black', lwd=bl_lwd, lty=3)

lines(ages, cas_sens1_part$final[1,], col='grey', lwd=gr_lwd)
lines(ages, cas_sens1_part$final[2,], col='grey', lwd=gr_lwd, lty=3)

for (c in 1:num_C2_models)
{
    lines(ages, c2_mat[c,1,], col=C2_color[c], lwd=1)
    lines(ages, c2_mat[c,2,], col=C2_color[c], lwd=1, lty=3)
}

plot_legend()
```


```{r plots_set_2, echo=FALSE}
# plot YCS

max_val <- max(c1_quant$`Vector parameter values`$recruitment.YCS,
               c1_sens1_quant$`Vector parameter values`$recruitment.YCS,
               max(unlist(list.map(cas2_mpd, max(Recruitment$ycs_values)))))

plot(c1_quant$`Vector parameter values`$recruitment.YCS, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='', main='YCS comparison')
lines(c1_sens1_quant$`Vector parameter values`$recruitment.YCS, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Recruitment$ycs_values, col=C2_color[c], lwd=1)
}

plot_legend()



# plot true YCS

max_val <- max(c1_quant$true_YCS$true_YCS,
               c1_sens1_quant$true_YCS$true_YCS,
               max(unlist(list.map(cas2_mpd, max(Recruitment$true_ycs)))))

plot(c1_quant$true_YCS$year, c1_quant$true_YCS$true_YCS, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='', main='true YCS comparison')
lines(c1_sens1_quant$true_YCS$year, c1_sens1_quant$true_YCS$true_YCS, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Recruitment$ycs_years, cas2_mpd[[c]]$Recruitment$true_ycs, col=C2_color[c], lwd=1)
}

plot_legend()



# plot recruitment

max_val <- max(c1_quant$recruitments$recruitment,
               c1_sens1_quant$recruitments$recruitment,
               max(unlist(list.map(cas2_mpd, max(Recruitment$Recruits)))))

plot(c1_quant$recruitments$year, c1_quant$recruitments$recruitment, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='', main='Recruitment comparison')
lines(c1_sens1_quant$recruitments$year, c1_sens1_quant$recruitments$recruitment, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines((cas2_mpd[[c]]$Recruitment$ycs_years + cas2_mpd[[c]]$Recruitment$ssb_offset), cas2_mpd[[c]]$Recruitment$Recruits, col=C2_color[c], lwd=1)
}

plot_legend()
```


```{r plots_set_3, echo=FALSE}
# plot true natural mortality-at-age

max_val <- max(c1_quant$`Ogive parameter values`$natural_mortality.ogive_all,
               c1_sens1_quant$`Ogive parameter values`$natural_mortality.ogive_all,
               max(unlist(list.map(cas2_mpd, max(NaturalMortalityOgive_all$Values)))))

plot(c1_quant$`Ogive parameter values`$natural_mortality.ogive_all, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Age', ylab='', main='Natural mortality-at-age comparison')
lines(c1_sens1_quant$`Ogive parameter values`$natural_mortality.ogive_all, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$NaturalMortalityOgive_all$Values, col=C2_color[c], lwd=1)
}

plot_legend()
```


```{r plots_set_4, echo=FALSE}
# plot summer survey
c1_surv      <- cas_mpd$free$`q[tan_sum].q` * unlist(c1_quant$Tangaroa_bio_summer_est)
c1_s1_surv   <- cas_mpd_sens1$free$`q[tan_sum].q` * unlist(c1_sens1_quant$Tangaroa_bio_summer_est)

c2_surv <- list()
for (c in 1:num_C2_models)
{
    c2_surv[[c]] <- cas2_mpd[[c]]$Tangaroa_bio_summer$Values
}

max_val <- max(c1_surv,
               c1_s1_surv,
               (c2_surv[[1]]$observed * exp(1.96 * c2_surv[[1]]$error_value)),
               max(unlist(list.map(c2_surv, max(expected)))))

plot(names(c1_surv), c1_surv, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='Biomass (t)', main='Tangaroa summer survey comparison')
lines(names(c1_s1_surv), c1_s1_surv, type='l', col='grey', lwd=gr_lwd)
points(c2_surv[[1]]$year, c2_surv[[1]]$observed, pch=20, col='black')
arrows(c2_surv[[1]]$year, (c2_surv[[1]]$observed * exp(-1.96 * c2_surv[[1]]$error_value)), c2_surv[[1]]$year, (c2_surv[[1]]$observed * exp(1.96 * c2_surv[[1]]$error_value)), length=0.05, angle=90, code=3)
for (c in 1:num_C2_models)
{
    points(c2_surv[[c]]$year, c2_surv[[c]]$expected, col=C2_color[c], pch=(15+c-1))
}

plot_legend()



# plot autumn survey

c1_surv      <- cas_mpd$free$`q[tan_aut].q` * unlist(c1_quant$Tangaroa_bio_autumn_est)
c1_s1_surv   <- cas_mpd_sens1$free$`q[tan_aut].q` * unlist(c1_sens1_quant$Tangaroa_bio_autumn_est)
c2_surv <- list()
for (c in 1:num_C2_models)
{
    c2_surv[[c]] <- cas2_mpd[[c]]$Tangaroa_bio_autumn$Values
}

max_val <- max(c1_surv,
               c1_s1_surv,
               (c2_surv[[1]]$observed * exp(1.96 * c2_surv[[1]]$error_value)),
               max(unlist(list.map(c2_surv, max(expected)))))

plot(names(c1_surv), c1_surv, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='Biomass (t)', main='Tangaroa autumn survey comparison')
lines(names(c1_s1_surv), c1_s1_surv, type='l', col='grey', lwd=gr_lwd)
points(c2_surv[[1]]$year, c2_surv[[1]]$observed, pch=20, col='black')
arrows(c2_surv[[1]]$year, (c2_surv[[1]]$observed * exp(-1.96 * c2_surv[[1]]$error_value)), c2_surv[[1]]$year, (c2_surv[[1]]$observed * exp(1.96 * c2_surv[[1]]$error_value)), length=0.05, angle=90, code=3)
for (c in 1:num_C2_models)
{
    points(c2_surv[[c]]$year, c2_surv[[c]]$expected, col=C2_color[c], pch=(15+c-1))
}

plot_legend()
```


```{r plots_set_5, echo=FALSE}
# > names(cas2_mpd_base$Mortality)
#  [1] "categories"                         "label"
#  [3] "m"                                  "selectivities"
#  [5] "time_step_ratio"                    "type"
#  [7] "year"                               "fishing_pressure[FishingLineHome]"
#  [9] "catch[FishingLineHome]"             "actual_catch[FishingLineHome]"
# [11] "fishing_pressure[FishingLineSpawn]" "catch[FishingLineSpawn]"
# [13] "actual_catch[FishingLineSpawn]"     "fishing_pressure[FishingTrwl]"
# [15] "catch[FishingTrwl]"                 "actual_catch[FishingTrwl]"

# plot fishing pressures - trawl

max_val <- max(c1_quant$fishing_pressures$trawl,
               c1_sens1_quant$fishing_pressures$trawl,
               max(unlist(list.map(cas2_mpd, max(Mortality$`fishing_pressure[FishingTrwl]`)))))

plot(c1_quant$fishing_pressures$year, c1_quant$fishing_pressures$trawl, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='U', main='Fishing pressure comparison: trawl', sub='Dotted line is summed Casal2 trawl and line_home')
lines(c1_sens1_quant$fishing_pressures$year, c1_sens1_quant$fishing_pressures$trawl, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Mortality$year, cas2_mpd[[c]]$Mortality$`fishing_pressure[FishingTrwl]`, col=C2_color[c], lwd=1)

    # sum fishing pressure for time step 1 to match CASAL output
    cas2_mpd[[c]]$Mortality$fishing_pressure_all <- cas2_mpd[[c]]$Mortality$`fishing_pressure[FishingTrwl]` +
                                                    cas2_mpd[[c]]$Mortality$`fishing_pressure[FishingLineHome]`

    lines(cas2_mpd[[c]]$Mortality$year, cas2_mpd[[c]]$Mortality$fishing_pressure_all, col=C2_color[c], lwd=1.5, lty=3)
}

plot_legend()



# plot fishing pressures - line_home

max_val <- max(c1_quant$fishing_pressures$line_home,
               c1_sens1_quant$fishing_pressures$line_home,
               max(unlist(list.map(cas2_mpd, max(Mortality$`fishing_pressure[FishingLineHome]`)))))

plot(c1_quant$fishing_pressures$year, c1_quant$fishing_pressures$line_home, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='U', main='Fishing pressure comparison: line_home', sub='Dotted line is summed Casal2 trawl and line_home')
lines(c1_sens1_quant$fishing_pressures$year, c1_sens1_quant$fishing_pressures$line_home, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Mortality$year, cas2_mpd[[c]]$Mortality$`fishing_pressure[FishingLineHome]`, col=C2_color[c], lwd=1)

    lines(cas2_mpd[[c]]$Mortality$year, cas2_mpd[[c]]$Mortality$fishing_pressure_all, col=C2_color[c], lwd=1.5, lty=3)
}

plot_legend()



# plot fishing pressures - line_spawn

max_val <- max(c1_quant$fishing_pressures$line_spawn,
               c1_sens1_quant$fishing_pressures$line_spawn,
               max(unlist(list.map(cas2_mpd, max(Mortality$`fishing_pressure[FishingLineSpawn]`)))))

plot(c1_quant$fishing_pressures$year, c1_quant$fishing_pressures$line_spawn, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='U', main='Fishing pressure comparison: line_spawn')
lines(c1_sens1_quant$fishing_pressures$year, c1_sens1_quant$fishing_pressures$line_spawn, col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Mortality$year, cas2_mpd[[c]]$Mortality$`fishing_pressure[FishingLineSpawn]`, col=C2_color[c], lwd=1)
}

plot_legend()
```


```{r plots_set_6, echo=FALSE}
ages <- seq(3, 25, 1)

max_val <- max(1,
               c1_quant$`Ogive parameter values`$`selectivity[surveysum_sel].male`, c1_quant$`Ogive parameter values`$`selectivity[surveysum_sel].female`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[surveysum_sel].male`, c1_sens1_quant$`Ogive parameter values`$`selectivity[surveysum_sel].female`)

plot(ages, c1_quant$`Ogive parameter values`$`selectivity[surveysum_sel].male`, type='l', col='black', lwd=bl_lwd, ylim=c(0,max_val), xlab='Age', ylab='', main='Summer survey selectivity-at-age comparison (male and female)', sub='Dotted lines are for female')
lines(ages, c1_quant$`Ogive parameter values`$`selectivity[surveysum_sel].female`, col='black', lwd=bl_lwd, lty=3)

lines(ages, c1_sens1_quant$`Ogive parameter values`$`selectivity[surveysum_sel].male`, col='grey', lwd=gr_lwd)
lines(ages, c1_sens1_quant$`Ogive parameter values`$`selectivity[surveysum_sel].female`, col='grey', lwd=gr_lwd, lty=3)

for (c in 1:num_C2_models)
{
    lines(ages, cas2_mpd[[c]]$summerTANSel_male$Values, col=C2_color[c], lwd=1)
    lines(ages, cas2_mpd[[c]]$summerTANSel_female$Values, col=C2_color[c], lwd=1, lty=3)
}

plot_legend()



max_val <- max(1,
               c1_quant$`Ogive parameter values`$`selectivity[surveyaut_sel].male`, c1_quant$`Ogive parameter values`$`selectivity[surveyaut_sel].female`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[surveyaut_sel].male`, c1_sens1_quant$`Ogive parameter values`$`selectivity[surveyaut_sel].female`)

plot(ages, c1_quant$`Ogive parameter values`$`selectivity[surveyaut_sel].male`, type='l', col='black', lwd=bl_lwd, ylim=c(0,max_val), xlab='Age', ylab='', main='Autumn survey selectivity-at-age comparison (male and female)', sub='Dotted lines are for female')
lines(ages, c1_quant$`Ogive parameter values`$`selectivity[surveyaut_sel].female`, col='black', lwd=bl_lwd, lty=3)

lines(ages, c1_sens1_quant$`Ogive parameter values`$`selectivity[surveyaut_sel].male`, col='grey', lwd=gr_lwd)
lines(ages, c1_sens1_quant$`Ogive parameter values`$`selectivity[surveyaut_sel].female`, col='grey', lwd=gr_lwd, lty=3)

for (c in 1:num_C2_models)
{
    lines(ages, cas2_mpd[[c]]$autumnTANSel_male$Values, col=C2_color[c], lwd=1)
    lines(ages, cas2_mpd[[c]]$autumnTANSel_female$Values, col=C2_color[c], lwd=1, lty=3)
}

plot_legend()



max_val <- max(1,
               c1_quant$`Ogive parameter values`$`selectivity[trawl_sel].male`, c1_quant$`Ogive parameter values`$`selectivity[trawl_sel].female`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[trawl_sel].male`, c1_sens1_quant$`Ogive parameter values`$`selectivity[trawl_sel].female`)

plot(ages, c1_quant$`Ogive parameter values`$`selectivity[trawl_sel].male`, type='l', col='black', lwd=bl_lwd, ylim=c(0,max_val), xlab='Age',  ylab='',main='Trawl fishery selectivity-at-age comparison (male and female)', sub='Dotted lines are for female')
lines(ages, c1_quant$`Ogive parameter values`$`selectivity[trawl_sel].female`, col='black', lwd=bl_lwd, lty=3)

lines(ages, c1_sens1_quant$`Ogive parameter values`$`selectivity[trawl_sel].male`, col='grey', lwd=gr_lwd)
lines(ages, c1_sens1_quant$`Ogive parameter values`$`selectivity[trawl_sel].female`, col='grey', lwd=gr_lwd, lty=3)

for (c in 1:num_C2_models)
{
    lines(ages, cas2_mpd[[c]]$trwlFSel_male$Values, col=C2_color[c], lwd=1)
    lines(ages, cas2_mpd[[c]]$trwlFSel_female$Values, col=C2_color[c], lwd=1, lty=3)
}

plot_legend()



max_val <- max(1,
               c1_quant$`Ogive parameter values`$`selectivity[line_home_sel].male`, c1_quant$`Ogive parameter values`$`selectivity[line_home_sel].female`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[line_home_sel].male`, c1_sens1_quant$`Ogive parameter values`$`selectivity[line_home_sel].female`)

plot(ages, c1_quant$`Ogive parameter values`$`selectivity[line_home_sel].male`, type='l', col='black', lwd=bl_lwd, ylim=c(0,max_val), xlab='Age',  ylab='',main='Longline home fishery selectivity-at-age comparison (male and female)', sub='Dotted lines are for female')
lines(ages, c1_quant$`Ogive parameter values`$`selectivity[line_home_sel].female`, col='black', lwd=bl_lwd, lty=3)

lines(ages, c1_sens1_quant$`Ogive parameter values`$`selectivity[line_home_sel].male`, col='grey', lwd=gr_lwd)
lines(ages, c1_sens1_quant$`Ogive parameter values`$`selectivity[line_home_sel].female`, col='grey', lwd=gr_lwd, lty=3)

for (c in 1:num_C2_models)
{
    lines(ages, cas2_mpd[[c]]$lineHomeFSel_male$Values, col=C2_color[c], lwd=1)
    lines(ages, cas2_mpd[[c]]$lineHomeFSel_female$Values, col=C2_color[c], lwd=1, lty=3)
}

plot_legend()



max_val <- max(1,
               c1_quant$`Ogive parameter values`$`selectivity[line_spawn_sel].male`, c1_quant$`Ogive parameter values`$`selectivity[line_spawn_sel].female`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[line_spawn_sel].male`, c1_sens1_quant$`Ogive parameter values`$`selectivity[line_spawn_sel].female`)

plot(ages, c1_quant$`Ogive parameter values`$`selectivity[line_spawn_sel].male`, type='l', col='black', lwd=bl_lwd, ylim=c(0,max_val), xlab='Age',  ylab='',main='Longline spawn fishery selectivity-at-age comparison (male and female)', sub='Dotted lines are for female')
lines(ages, c1_quant$`Ogive parameter values`$`selectivity[line_spawn_sel].female`, col='black', lwd=bl_lwd, lty=3)

lines(ages, c1_sens1_quant$`Ogive parameter values`$`selectivity[line_spawn_sel].male`, col='grey', lwd=gr_lwd)
lines(ages, c1_sens1_quant$`Ogive parameter values`$`selectivity[line_spawn_sel].female`, col='grey', lwd=gr_lwd, lty=3)

for (c in 1:num_C2_models)
{
    lines(ages, cas2_mpd[[c]]$lineSpawnFSel_male$Values, col=C2_color[c], lwd=1)
    lines(ages, cas2_mpd[[c]]$lineSpawnFSel_female$Values, col=C2_color[c], lwd=1, lty=3)
}

plot_legend()
```

