---
title: "Casal2 Test Case comparisons for HAK"
output:
    html_document:
        toc: TRUE
        toc_depth: 2
        number_sections: TRUE

header-includes:
 - \usepackage{pdflscape}
 - \newcommand{\blandscape}{\begin{landscape}}
 - \newcommand{\elandscape}{\end{landscape}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# HAK comparison of CASAL and Casal2 model configurations

This document compares the results of at least 2 CASAL model configurations (base and at least one sensitivity) and up to 8 Casal2 model configurations (3 BetaDiff, 2 CppAD, and 3 ADOL-C).

The CASAL model sensitivity 1 has a smaller tolerance value than the CASAL base model (1e-6 vs. 2e-3).

The Casal2 ADOL-C and BetaDiff low tolerance models have a smaller tolerance value than the CASAL base model (1e-6 vs. 2e-3). The Casal2 CppAD models have a tolerance value of 1e-9.


## HAK model characteristics

The main characteristics of the Test Case HAK (hake) CASAL model are:

* one stock, ages 1 - 30
* one area
* years 1975 - 2018, projection years 2019 - 2022
* two time steps, duration 7 months and 5 months
* recruitment, spawning, aging, maturation, fishing in time step 1
* two stock categories: males and females, with user-provided maturity-at-age ogive
* natural mortality (M) of 0.19
* Beverton-Holt stock-recruitment relationship, with steepness (h) set to 0.8
* sex-specific Schnute age-length relationship
* sex-specific length-weight relationship ($W = aL^b$)
* ageing error type normal (c = 0.08)
* one fishery with logistic selectivity-at-age
* one survey with logistic selectivity-at-age

Observation data include:

* fishery CPUE index
* fishery proportions-at-age
* survey biomass index
* survey proportions-at-age

Parameters estimated include:

* B0
* YCS
* catchability (q) for the fishery CPUE index and for the survey index
* fishery and survey selectivity parameters a50 and ato95


## R environment

```{r C1_C2_setup, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_set_up_R_env.R')
```


## CASAL and Casal2 model output

```{r CASAL, warning=FALSE}
source('../../R-functions/report_read_in_CASAL_MPD_files.R')
```


```{r Casal2}
source('../../R-functions/report_read_in_Casal2_MPD_files.R')
```



\newpage
\blandscape
## Tables

Tables of parameter estimates and objective function components for the CASAL and Casal2 model MPD results

```{r table_setup, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_set_up_tables.R')

options(scipen=999)
```


```{r huxtables_parameters, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_print_parameter_tables.R')

hux_C1_est_params
hux_C2_est_params
hux_C2_pd_est_params
```


```{r huxtables_obj_fun, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_print_obj_fun_tables.R')

hux_C1_obj_fun
hux_C2_obj_fun
```


```{r convergence_info, echo=FALSE}
source('../../R-functions/report_print_convergence_info.R')
```


```{r warnings_and_bounds, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_print_warnings.R')
```


```{r huxtables_close, echo=FALSE}
options(scipen=0)
```
\elandscape
\newpage



## Matching of outputs

Time series comparisons with CASAL base model results

```{r Time_Series_Match, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_time_series_match_function.R')


for (c in 1:num_C2_models)
{
    print(paste('Catch time series base model comparison for run', C2_subdir[c]))

    print(paste('Actual catches for fishery match:',
                time_series_match(c1_quant$actual_catches$catch, cas2_mpd[[c]]$Mortality$`actual_catch[subaF]`)))

    print('')
}
```

Derived quantities

SB0, SBcurrent, MSY, F_MSY, others...



## Plots

Comparison plots

```{r plots_base, echo=FALSE}

# c(bottom, left, top, right)
par(mar=c(4,4,2,1) + 0.1)

par(mfrow=c(1,2))
```


```{r plots_correlation, echo=FALSE}
source('../../R-functions/report_plot_correlation_matrices.R')
```


```{r plots_set_0, echo=FALSE}
source('../../R-functions/report_set_up_legend.R')


max_val <- max(c1_quant$SSBs$SSB,
               c1_sens1_quant$SSBs$SSB,
               max(unlist(list.map(cas2_mpd, max(DerivedQuantities$SSB$values)))))

plot(c1_quant$SSBs$year, c1_quant$SSBs$SSB, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='Biomass (t)', main='SSB comparison')
lines(c1_sens1_quant$SSBs$year, c1_sens1_quant$SSBs$SSB, type='l', col='grey', lwd=2.5)

for (c in 1:num_C2_models)
{
    lines(names(cas2_mpd[[c]]$DerivedQuantities$SSB$values), cas2_mpd[[c]]$DerivedQuantities$SSB$values, col=C2_color[c], lwd=1)
}

plot_legend()
```


```{r plots_set_1, echo=FALSE}

# Q:  how to get CASAL initial age structure?
# A:  use 'casal::extract.partition()''


ages <- seq(1, 30)

# omit the 'category' column
c2_mat <- array(0, dim=c(num_C2_models, dim(as.matrix(cas2_mpd[[1]]$Init$values[,-1]))))
for (c in 1:num_C2_models)
{
    c2_mat[c,,] <- as.matrix(cas2_mpd[[c]]$Init$values[,-1])
}

# plot initial numbers

max_val <- max(cas_part$initial,
               cas_sens1_part$initial,
               max(c2_mat))

plot(ages, cas_part$initial[1,], type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Age', ylab='', main='Initial numbers-at-age comparison')
lines(ages, cas_part$initial[2,], col='black', lwd=5, lty=3)
lines(ages, cas_sens1_part$initial[1,], col='grey', lwd=2.5)
lines(ages, cas_sens1_part$initial[2,], col='grey', lwd=2.5, lty=3)
for (c in 1:num_C2_models)
{
    lines(ages, c2_mat[c,1,], col=C2_color[c], lwd=1)
    lines(ages, c2_mat[c,2,], col=C2_color[c], lwd=1, lty=3)
}

plot_legend()



c2_mat <- array(0, dim=c(num_C2_models, dim(as.matrix(cas2_mpd[[1]]$state1$`2018`$values))))
for (c in 1:num_C2_models)
{
    c2_mat[c,,] <- as.matrix(cas2_mpd[[c]]$state1$`2018`$values)
}

# plot final year numbers (time step 1)

max_val <- max(cas_part$final,
               cas_sens1_part$final,
               max(c2_mat))

plot(ages, cas_part$final[1,], type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Age', ylab='', main='Final numbers-at-age comparison (time step 1)')
lines(ages, cas_part$final[2,], col='black', lwd=5, lty=3)
lines(ages, cas_sens1_part$final[1,], col='grey', lwd=2.5)
lines(ages, cas_sens1_part$final[2,], col='grey', lwd=2.5, lty=3)
for (c in 1:num_C2_models)
{
    lines(ages, c2_mat[c,1,], col=C2_color[c], lwd=1)
    lines(ages, c2_mat[c,2,], col=C2_color[c], lwd=1, lty=3)
}

plot_legend()



c2_mat <- array(0, dim=c(num_C2_models, dim(as.matrix(cas2_mpd[[1]]$state2$`2018`$values))))
for (c in 1:num_C2_models)
{
    c2_mat[c,,] <- as.matrix(cas2_mpd[[c]]$state2$`2018`$values)
}

# plot final year numbers (time step 2)

max_val <- max(cas_part$final,
               cas_sens1_part$final,
               max(c2_mat))

plot(ages, cas_part$final[1,], type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Age', ylab='', main='Final numbers-at-age comparison (time step 2)')
lines(ages, cas_part$final[2,], col='black', lwd=5, lty=3)
lines(ages, cas_sens1_part$final[1,], col='grey', lwd=2.5)
lines(ages, cas_sens1_part$final[2,], col='grey', lwd=2.5, lty=3)
for (c in 1:num_C2_models)
{
    lines(ages, c2_mat[c,1,], col=C2_color[c], lwd=1)
    lines(ages, c2_mat[c,2,], col=C2_color[c], lwd=1, lty=3)
}

plot_legend()
```


```{r plots_set_2, echo=FALSE}
# plot YCS

max_val <- max(c1_quant$`Vector parameter values`$recruitment.YCS,
               c1_sens1_quant$`Vector parameter values`$recruitment.YCS,
               max(unlist(list.map(cas2_mpd, max(Recruitment$ycs_values)))))

plot(c1_quant$`Vector parameter values`$recruitment.YCS, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='', main='YCS comparison')
lines(c1_sens1_quant$`Vector parameter values`$recruitment.YCS, type='l', col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Recruitment$ycs_values, col=C2_color[c], lwd=1)
}

plot_legend()



# plot true YCS

max_val <- max(c1_quant$true_YCS$true_YCS,
               c1_sens1_quant$true_YCS$true_YCS,
               max(unlist(list.map(cas2_mpd, max(Recruitment$true_ycs)))))

plot(c1_quant$true_YCS$year, c1_quant$true_YCS$true_YCS, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='', main='true YCS comparison')
lines(c1_sens1_quant$true_YCS$year, c1_sens1_quant$true_YCS$true_YCS, type='l', col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Recruitment$ycs_years, cas2_mpd[[c]]$Recruitment$true_ycs, col=C2_color[c], lwd=1)
}

plot_legend()



# plot recruitment

max_val <- max(c1_quant$recruitments$recruitment,
               c1_sens1_quant$recruitments$recruitment,
               max(unlist(list.map(cas2_mpd, max(Recruitment$Recruits)))))

plot(c1_quant$recruitments$year, c1_quant$recruitments$recruitment, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='', main='Recruitment comparison')
lines(c1_sens1_quant$recruitments$year, c1_sens1_quant$recruitments$recruitment, type='l', col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Recruitment$ycs_years, cas2_mpd[[c]]$Recruitment$Recruits, col=C2_color[c], lwd=1)
}

plot_legend()
```


```{r plots_set_3, echo=FALSE}
# plot survey
c1_surv    <- unlist(c1_quant$subaTANbiomassDEC.fits)
c1_s1_surv <- unlist(c1_sens1_quant$subaTANbiomassDEC.fits)

c2_surv <- list()
for (c in 1:num_C2_models)
{
    c2_surv[[c]] <- cas2_mpd[[c]]$Tangaroa_biomass$Values
}

max_val <- max(c1_surv,
               c1_s1_surv,
               (c2_surv[[1]]$observed * exp(1.96 * c2_surv[[1]]$error_value)),
               max(unlist(list.map(c2_surv, max(expected)))))

plot(names(c1_surv), c1_surv, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='Biomass (t)', main='Tangaroa summer survey comparison')
lines(names(c1_s1_surv), c1_s1_surv, type='l', col='grey', lwd=2.5)
points(c2_surv[[1]]$year, c2_surv[[1]]$observed, pch=20, col='black')
arrows(c2_surv[[1]]$year, (c2_surv[[1]]$observed * exp(-1.96 * c2_surv[[1]]$error_value)), c2_surv[[1]]$year, (c2_surv[[1]]$observed * exp(1.96 * c2_surv[[1]]$error_value)), length=0.05, angle=90, code=3)
for (c in 1:num_C2_models)
{
    points(c2_surv[[c]]$year, c2_surv[[c]]$expected, col=C2_color[c], pch=(15+c-1))
}

plot_legend()
```


```{r plots_set_4, echo=FALSE}
# plot fishery CPUE
c1_surv    <- unlist(c1_quant$CPUE.fits)
c1_s1_surv <- unlist(c1_sens1_quant$CPUE.fits)

c2_surv <- list()
for (c in 1:num_C2_models)
{
    c2_surv[[c]] <- cas2_mpd[[c]]$CPUE_index$Values
}

max_val <- max(c1_surv,
               c1_s1_surv,
               (c2_surv[[1]]$observed * exp(1.96 * c2_surv[[1]]$error_value)),
               max(unlist(list.map(c2_surv, max(expected)))))

plot(names(c1_surv), c1_surv, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='Biomass (t)', main='Fishery CPUE index comparison', sub='Compare process error estimates')
lines(names(c1_s1_surv), c1_s1_surv, type='l', col='grey', lwd=2.5)
points(c2_surv[[1]]$year, c2_surv[[1]]$observed, pch=20, col='black')
arrows(c2_surv[[1]]$year, (c2_surv[[1]]$observed * exp(-1.96 * c2_surv[[1]]$error_value)), c2_surv[[1]]$year, (c2_surv[[1]]$observed * exp(1.96 * c2_surv[[1]]$error_value)), length=0.05, angle=90, code=3)
for (c in 1:num_C2_models)
{
    points(c2_surv[[c]]$year, c2_surv[[c]]$expected, col=C2_color[c], pch=(15+c-1))
}

plot_legend()
```


```{r plots_set_5, echo=FALSE}
# > names(cas2_mpd[[1]])
#  [1] "Init"                   "state1"                 "state2"
#  [4] "obj_fun"                "DQs"                    "estimated_values"
#  [7] "estimated_summary"      "Qs"                     "SSB"
# [10] "Mortality"              "Recruitment"            "maturity_ogive_m"
# [13] "maturity_ogive_f"       "growth_length_at_age_1" "growth_length_at_age_2"
# [16] "growth_weight_at_age_1" "growth_weight_at_age_2" "fishery_selectivity"
# [19] "survey_selectivity"     "CPUE_index"             "fishery_AF"
# [22] "Tangaroa_biomass"       "Tangaroa_AF"            "Hess"
# [25] "Covar"                  "Corr"                   "warnings_encounted"

# plot fishing pressures

max_val <- max(c1_quant$fishing_pressures$pressure,
               c1_sens1_quant$fishing_pressures$pressure,
               max(unlist(list.map(cas2_mpd, max(Mortality$`fishing_pressure[subaF]`)))))

plot(c1_quant$fishing_pressures$year, c1_quant$fishing_pressures$pressure, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='U', main='Fishing pressure comparison')
lines(c1_sens1_quant$fishing_pressures$year, c1_sens1_quant$fishing_pressures$pressure, type='l', col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Mortality$year, cas2_mpd[[c]]$Mortality$`fishing_pressure[subaF]`, col=C2_color[c], lwd=1)
}

plot_legend()
```


```{r plots_set_6, echo=FALSE}
ages <- seq(1, 30)

max_val <- max(1,
               c1_quant$`Ogive parameter values`$`selectivity[subaFsel].all`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[subaFsel].all`)

plot(ages, c1_quant$`Ogive parameter values`$`selectivity[subaFsel].all`, type='l', col='black', lwd=5, ylim=c(0,max_val), xlab='Age', ylab='', main='Fishery selectivity-at-age comparison')
lines(ages, c1_sens1_quant$`Ogive parameter values`$`selectivity[subaFsel].all`, col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(ages, cas2_mpd[[c]]$fishery_selectivity$Values, col=C2_color[c], lwd=1)
}

plot_legend()
```


```{r plots_set_7, echo=FALSE}
ages <- seq(1, 30)

max_val <- max(1,
               c1_quant$`Ogive parameter values`$`selectivity[subaTANselA].all`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[subaTANselA].all`)

plot(ages, c1_quant$`Ogive parameter values`$`selectivity[subaTANselA].all`, type='l', col='black', lwd=5, ylim=c(0,max_val), xlab='Age', ylab='', main='Survey selectivity-at-age comparison')
lines(ages, c1_sens1_quant$`Ogive parameter values`$`selectivity[subaTANselA].all`, col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(ages, cas2_mpd[[c]]$survey_selectivity$Values, col=C2_color[c], lwd=1)
}

plot_legend()
```

