---
title: "Casal2 Test Case MCMC comparisons for HAK"
output:
    html_document:
        toc: TRUE
        toc_depth: 2
        number_sections: TRUE

header-includes:
 - \usepackage{pdflscape}
 - \newcommand{\blandscape}{\begin{landscape}}
 - \newcommand{\elandscape}{\end{landscape}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# HAK comparison of CASAL and Casal2 model configurations

This document compares the results of at least 2 CASAL model configurations (base and at least one sensitivity) and up to 8 Casal2 model configurations (3 BetaDiff, 2 CppAD, and 3 ADOL-C).

The CASAL model sensitivity 1 has a smaller minimisation tolerance value than the CASAL base model (1e-6 vs. 2e-3).

The Casal2 ADOL-C and BetaDiff low tolerance models have a smaller tolerance value than the CASAL base model (1e-6 vs. 2e-3). The Casal2 CppAD models have a tolerance value of 1e-9.


## HAK model characteristics

The main characteristics of the Test Case HAK (hake) CASAL model are:

* one stock, ages 1 - 30
* one area
* years 1975 - 2018, projection years 2019 - 2022
* two time steps, duration 7 months and 5 months
* recruitment, spawning, aging, maturation, fishing in time step 1
* two stock categories: males and females, with user-provided maturity-at-age ogive
* natural mortality (M) of 0.19
* Beverton-Holt stock-recruitment relationship, with steepness (h) 0.8
* sex-specific Schnute age-length relationship
* sex-specific length-weight relationship ($W = aL^b$)
* ageing error type normal (c = 0.08)
* one fishery with logistic selectivity-at-age
* one survey with logistic selectivity-at-age

Observation data include:

* fishery CPUE index
* fishery proportions-at-age
* survey biomass index
* survey proportions-at-age

Parameters estimated include:

* B0
* catchability (q) for the fishery CPUE index and for the survey index
* fishery and survey selectivity parameters a50 and ato95
* YCS (recruitment deviations)


## R environment

```{r C1_C2_setup, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_set_up_R_env.R')
```


## CASAL and Casal2 model output

```{r CASAL, warning=FALSE}
source('../../R-functions/report_read_in_CASAL_MPD_files.R')
source('../../R-functions/report_read_in_CASAL_MCMC_files.R')
```


```{r Casal2}
source('../../R-functions/report_read_in_Casal2_MPD_files.R')
source('../../R-functions/report_read_in_Casal2_MCMC_files.R')
```


## MCMC diagnostics

Diagnostics for the CASAL and Casal2 model MCMC results

```{r CASAL_MCMC_diag, message=FALSE}
num_samples_to_use <- 10000
MCMC_params_rename <- c('objective_function', 'SB0', 'q_CPUE', 'q_subaTANqA')

C1_MCMC_params <- c('neg_ln_posterior', 'initialization.B0', 'q[cpue].q', 'q[subaTANqA].q')

source('../../R-functions/report_CASAL_MCMC_diagnostics.R')
```


```{r Casal2_MCMC_diag, message=FALSE}
C2_MCMC_params <- c('objective_score', 'process[Recruitment].b0', 'catchability[CPUEq].q', 'catchability[subaTANqA].q')

source('../../R-functions/report_Casal2_MCMC_diagnostics.R')
```

```{r plot_MCMC_diag, message=FALSE}
source('../../R-functions/report_plot_MCMC_diagnostics.R')
```


