---
title: "Casal2 Test Case comparisons for ORH"
output:
    html_document:
        toc: true
        toc_depth: 2
        number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ORH comparison of CASAL and Casal2 model configurations

This document compares the results of at least 2 CASAL model configurations (base and at least one sensitivity) and up to 8 Casal2 model configurations (3 BetaDiff, 2 CppAD, and 3 ADOL-C).

The CASAL model sensitivity 1 has a smaller tolerance value than the CASAL base model (1e-6 vs. 2e-3).

The Casal2 ADOL-C and BetaDiff low tolerance models have a smaller tolerance value than the CASAL base model (1e-6 vs. 2e-3). The Casal2 CppAD models have a tolerance value of 1e-9.


## ORH model characteristics

The main characteristics of the Test Case ORH (orange roughy) CASAL model are:

* one stock, ages 1 - 80
* one area
* years 1979 - 2005, projection years 2006 - 2010
* one time step
* two stock categories: immature and mature, with logistic-producing maturity ogive
* natural mortality (M) of 0.045
* Beverton-Holt stock-recruitment relationship, with steepness (h) set to 1.0
* length-weight relationship ($W = aL^b$)
* no ageing error
* one fishery with logistic selectivity-at-age

Observation data include:

* empirical length-at-age data for one year
* one CPUE index
* fishery proportions-at-age

Parameters estimated include:

* B0
* the first and last CV for the age-length relationship
* nuisance catchability (q) for the fishery CPUE index
* fishery selectivity parameters a50 and ato95


## R environment

```{r C1_C2_setup, echo=FALSE, warning=FALSE, message=FALSE}
# print(paste('pandoc installed?', rmarkdown::pandoc_available()))
# print(paste('pandoc version:', rmarkdown::pandoc_version()))

require(ggplot2)
require(rlist)


require(devtools)

# the CASAL and Casal2 R libraries are installed in the TestCases subdirectory
.libPaths(c('../..', .libPaths()))

if (!require(casal))
{
    install.packages('../../../R-libraries/casal_2.30.tar.gz', repos=NULL)
    library(casal)
    # install('casal_utils')
}

if (!require(casal2))
{
    install.packages('../../../R-libraries/casal2_1.0.tar.gz', repos=NULL)
    library(casal2)
}

date()
sessionInfo()

base_dir <- '.'

# MPD files are params_est.out, run_estimation.txt, and mpd.out for both CASAL and Casal2
params_est <- 'params_est.out'
mpd_run    <- 'run_estimation.txt'
mpd_out    <- 'mpd.out'

get_convergence_information <- function(filename)
{
    # for BetaDiff
    # case-insensitive check for all strings with 'converge' and without 'check' (for "Convergence check")
    str_vec <- grep('check', grep('converge|fmm|optimise', readLines(filename), ignore.case=TRUE, value=TRUE), value=TRUE, invert=TRUE)

    # for ADOL-C
    if (length(str_vec) == 0)
    {
        # str_vec <- ?????
    }

    # for CppAD
    if (length(str_vec) == 0)
    {
        str_vec <- grep('Number of|EXIT', readLines(filename), value=TRUE)
    }

    return (str_vec)
}

```


## CASAL and Casal2 model output

```{r CASAL, warning=FALSE}
# read in CASAL MPD results
cas_mpd_filename <- file.path(base_dir, 'CASAL', mpd_run)
cas_mpd  <- casal::extract.mpd(cas_mpd_filename)
cas_part <- casal::extract.partition(cas_mpd_filename)
cas_corr <- casal::extract.correlation.matrix(cas_mpd_filename)
cas_conv <- get_convergence_information(cas_mpd_filename)
c1_quant <- cas_mpd$quantities

cas_mpd_sens1_filename <- file.path(base_dir, 'CASAL_sens1', mpd_run)
cas_mpd_sens1  <- casal::extract.mpd(cas_mpd_sens1_filename)
cas_sens1_part <- casal::extract.partition(cas_mpd_sens1_filename)
cas_sens1_corr <- casal::extract.correlation.matrix(cas_mpd_sens1_filename)
cas_sens1_conv <- get_convergence_information(cas_mpd_sens1_filename)
c1_sens1_quant <- cas_mpd_sens1$quantities
```


```{r Casal2}
C2_subdir <- c('betadiff_casal_flags_on',
               'betadiff_casal_flags_off',
               'betadiff_casal_flags_on_low_tol',
               'cppad_casal_flags_on',
               'cppad_casal_flags_off',
               'adolc_casal_flags_on',
               'adolc_casal_flags_off',
               'adolc_casal_flags_on_low_tol')


# find the non-zero-sized parameter estimate files for the C2 models
C2_subdir <- unlist(sapply(C2_subdir, function(c_dir) if ( file.access(file.path(base_dir, 'Casal2', c_dir, params_est), 4) == 0 &
                                                           file.size(file.path(base_dir, 'Casal2', c_dir, params_est)) > 0 )
                                                         { return (c_dir) }, USE.NAMES=FALSE ))


num_C2_models <- length(C2_subdir)

C2_color <- c('blue', 'green3', 'red', 'gold', 'magenta', 'cyan', 'brown', 'orange')

# read in Casal2 MPD results
cas2_mpd  <- list()
cas2_corr <- list()
cas2_conv <- list()
for (c in 1:num_C2_models)
{
    cas2_mpd_filename <- file.path(base_dir, 'Casal2', C2_subdir[c], mpd_run)
    cas2_mpd[[c]] <- casal2::extract.mpd(cas2_mpd_filename)

    cas2_corr[[c]] <- casal2::extract.correlation.matrix(mpd_out, file.path(base_dir, 'Casal2', C2_subdir[c]))

    cas2_conv[[c]] <- get_convergence_information(cas2_mpd_filename)
}
```



## Tables

Tables of parameter estimates and objective function components for the CASAL and Casal2 model MPD results

```{r table_setup, echo=FALSE, warning=FALSE, message=FALSE}
require(plyr)
require(dplyr)
require(huxtable)


# options(huxtable.knitr_output_format = "html")


rotate_and_label_cols <- function(df_1, labels)
{
    loc_df <- t(df_1)
    colnames(loc_df) <- labels

    return (loc_df)
}


reformat_to_huxtable <- function(df_1, caption_str=NA, format_num=3, reformat_col=FALSE, col_num=-1, add_rownames=TRUE)
{
    hux_table <- as_hux(df_1, add_colnames=TRUE, add_rownames=add_rownames)

    hux_nr <- nrow(hux_table)
    hux_nc <- ncol(hux_table)

    top_border(hux_table)[1,] <- 1
    bottom_border(hux_table)[c(1,hux_nr),] <- 1
    left_border(hux_table)[,1:2] <- 1
    right_border(hux_table)[,hux_nc] <- 1
    caption(hux_table) <- caption_str

    # reformat numbers in table except for the top row and left column
    if (format_num >= 0)
    {
        hux_table <- set_number_format(hux_table, -1, -1, format_num)
    }

    # to reformat and highlight specific columns
    align(hux_table)[-1,col_num] <- '.'
    if (reformat_col & col_num != 0)
    {
        hux_table <- set_number_format(hux_table, -1, col_num, 3)
        hux_table <- map_background_color(hux_table, -1, col_num,
                                          by_ranges(c(-20, -10, 0, 10, 20), c('orange', 'pink', 'white', 'white', 'cyan', 'yellow')))
    }

    return (hux_table)
}


C2_est_params <- cas2_mpd[[1]]$estimated_values$values
c2_est_params_1 <- C2_est_params
C2_obj_fun    <- cas2_mpd[[1]]$obj_fun$values

C2_pd_est_params <- C2_est_params - C2_est_params

if (num_C2_models > 1)
{
    for (c in 2:num_C2_models)
    {
        c2_est_params_c <- cas2_mpd[[c]]$estimated_values$values
        C2_est_params <- bind_rows(C2_est_params, c2_est_params_c)

        C2_pd_est_params <- bind_rows(C2_pd_est_params, 100.0 * (1.0 - (c2_est_params_c / c2_est_params_1)))

        C2_obj_fun    <- bind_rows(C2_obj_fun, cas2_mpd[[c]]$obj_fun$values)
    }
}

options(scipen=999)
```


```{r huxtables_parameters, echo=FALSE, warning=FALSE, message=FALSE}
C1_est_params <- cbind(unlist(cas_mpd$free), unlist(cas_mpd_sens1$free), 100.0 * (1.0 - (unlist(cas_mpd_sens1$free) / unlist(cas_mpd$free))))
colnames(C1_est_params) <- c('Base_Model', 'Sensitivity_1', 'Percent_Diff')
# print('CASAL parameter estimates')
# print(C1_est_params, digits=5)
hux_C1_est_params <- reformat_to_huxtable(C1_est_params, 'CASAL parameter estimates and Percent Difference', 5, TRUE, 4)
hux_C1_est_params


C2_est_params <- rotate_and_label_cols(C2_est_params, C2_subdir)
# print('Casal2 parameter estimates')
# print(C2_est_params, digits=5)
hux_C2_est_params <- reformat_to_huxtable(C2_est_params, 'Casal2 parameter estimates', 5)
hux_C2_est_params


C2_pd_est_params <- rotate_and_label_cols(C2_pd_est_params, C2_subdir)
# print('Casal2 parameter estimate percent differences')
# print(C2_pd_est_params, digits=3)
hux_C2_pd_est_params <- reformat_to_huxtable(C2_pd_est_params, paste('Casal2 parameter estimates: Percent Difference from ', C2_subdir[1], sep=''), 3, TRUE, -1)
hux_C2_pd_est_params
```

```{r huxtables_obj_fun, echo=FALSE, warning=FALSE, message=FALSE}
C1_obj_fun <- data.frame(Component=cas_mpd$objective.function$components$label,
                         Base_Model=cas_mpd$objective.function$components$value,
                         Sensitivity_1=cas_mpd_sens1$objective.function$components$value)
C1_obj_fun <- bind_rows(C1_obj_fun,
                        data.frame(Component='Total', Base_Model=cas_mpd$objective.function$value, Sensitivity_1=cas_mpd_sens1$objective.function$value))
# print('CASAL objective function values')
# print(C1_obj_fun, digits=5)
hux_C1_obj_fun <- reformat_to_huxtable(C1_obj_fun, 'CASAL objective function component values', add_rownames=FALSE)
hux_C1_obj_fun


C2_obj_fun <- rotate_and_label_cols(C2_obj_fun, C2_subdir)
# print('Casal2 objective function values')
# print(C2_obj_fun, digits=5)
hux_C2_obj_fun <- reformat_to_huxtable(C2_obj_fun, 'Casal2 objective function component values')
hux_C2_obj_fun
```


```{r convergence_info, echo=FALSE}
print('CASAL base model convergence information')
print(cas_conv)
print('')

print('CASAL sensitivity 1 model convergence information')
print(cas_sens1_conv)

for (c in 1:num_C2_models)
{
    print('')
    print(paste('Casal2', C2_subdir[c], 'model convergence information'))
    print(cas2_conv[[c]])
}
```


```{r warnings_and_bounds, echo=FALSE, warning=FALSE, message=FALSE}
print('CASAL base model warnings')
print(cas_mpd$parameters.at.bounds)
print('')

print('CASAL sensitivity 1 model warnings')
print(cas_mpd_sens1$parameters.at.bounds)

for (c in 1:num_C2_models)
{
    print('')
    print(paste('Casal2', C2_subdir[c], 'model warnings'))
    print(cas2_mpd[[c]]$warnings_encounted)
}
```


```{r huxtables_close, echo=FALSE}
options(scipen=0)
```



## Matching of outputs

Time series comparisons with CASAL base model results

```{r Time_Series_Match, echo=FALSE, warning=FALSE, message=FALSE}
# do the catch time series differ?

time_series_match <- function(t1_vec, t2_vec)
{
    if (length(t1_vec) == length(t2_vec))
    {
        return (ifelse(max(abs(range(t1_vec - t2_vec))) == 0, 'yes', 'no'))
    }

    return ('no')
}


for (c in 1:num_C2_models)
{
    print(paste('Catch time series base model comparison for run', C2_subdir[c]))

    print(paste('Actual catches for Andes match:',
                time_series_match(c1_quant$actual_catches$catch, cas2_mpd[[c]]$Mortality$`actual_catch[AndesFishery]`)))

    print('')
}
```

Derived quantities

SB0, SBcurrent, MSY, F_MSY, others...



## Plots

Comparison plots

```{r plots_base, echo=FALSE}

# c(bottom, left, top, right)
par(mar=c(4,4,2,1) + 0.1)

par(mfrow=c(1,2))
```


```{r plots_correlation, echo=FALSE}

plot_image_and_range <- function(corr_obj, label)
{
    image(corr_obj, zlim=c(-1,1), col=hcl.colors(51, 'RdYlBu'), xlab='Parameter', ylab='Parameter', main=label)
    print(paste(label, 'range (excluding 1.0):', paste0(range(corr_obj[corr_obj < 1], na.rm=TRUE), collapse=' ')))
}

plot_image_and_range(cas_corr, 'CASAL base parameter correlation')

plot_image_and_range(cas_sens1_corr, 'CASAL sensitivity 1 parameter correlation')

for (c in 1:num_C2_models)
{
    plot_image_and_range(cas2_corr[[c]], paste('Casal2', C2_subdir[c], 'parameter correlation'))
}
```


```{r plots_set_0, echo=FALSE}
plot_legend <- function()
{
    # placeholder plot for legend
    plot(0, 0, type='n', xlim=c(0,1), ylim=c(0,1), axes=FALSE, xlab='', ylab='', main='')
    legend(0.1, 0.99, c('CASAL', 'CASAL sens1', C2_subdir), lwd=5, col=c('black', 'grey', C2_color), pt.cex=1.6)
}


max_val <- max(c1_quant$SSBs$SSB,
               c1_sens1_quant$SSBs$SSB,
               max(unlist(list.map(cas2_mpd, max(SSB$SSB$values)))))

plot(c1_quant$SSBs$year, c1_quant$SSBs$SSB, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='Biomass (t)', main='SSB comparison')
lines(c1_sens1_quant$SSBs$year, c1_sens1_quant$SSBs$SSB, type='l', col='grey', lwd=2.5)

for (c in 1:num_C2_models)
{
    lines(names(cas2_mpd[[c]]$SSB$SSB$values), cas2_mpd[[c]]$SSB$SSB$values, col=C2_color[c], lwd=1)
}

plot_legend()
```


```{r plots_set_1, echo=FALSE}

# Q:  how to get CASAL initial age structure?
# A:  use 'casal::extract.partition()''


ages <- seq(1, 80)

# omit the 'category' column
c2_mat <- array(0, dim=c(num_C2_models, dim(as.matrix(cas2_mpd[[1]]$Init$values[,-1]))))
for (c in 1:num_C2_models)
{
    c2_mat[c,,] <- as.matrix(cas2_mpd[[c]]$Init$values[,-1])
}

# plot initial numbers

max_val <- max(cas_part$initial,
               cas_sens1_part$initial,
               max(c2_mat))

plot(ages, cas_part$initial[1,], type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Age', ylab='', main='Initial numbers-at-age comparison')
lines(ages, cas_part$initial[2,], col='black', lwd=5, lty=3)
lines(ages, cas_sens1_part$initial[1,], col='grey', lwd=2.5)
lines(ages, cas_sens1_part$initial[2,], col='grey', lwd=2.5, lty=3)
for (c in 1:num_C2_models)
{
    lines(ages, c2_mat[c,1,], col=C2_color[c], lwd=1)
    lines(ages, c2_mat[c,2,], col=C2_color[c], lwd=1, lty=3)
}

plot_legend()



c2_mat <- array(0, dim=c(num_C2_models, dim(as.matrix(cas2_mpd[[1]]$state1$`2005`$values))))
for (c in 1:num_C2_models)
{
    c2_mat[c,,] <- as.matrix(cas2_mpd[[c]]$state1$`2005`$values)
}

# plot final numbers

max_val <- max(cas_part$final,
               cas_sens1_part$final,
               max(c2_mat))

plot(ages, cas_part$final[1,], type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Age', ylab='', main='Final numbers-at-age comparison')
lines(ages, cas_part$final[2,], col='black', lwd=5, lty=3)
lines(ages, cas_sens1_part$final[1,], col='grey', lwd=2.5)
lines(ages, cas_sens1_part$final[2,], col='grey', lwd=2.5, lty=3)
for (c in 1:num_C2_models)
{
    lines(ages, c2_mat[c,1,], col=C2_color[c], lwd=1)
    lines(ages, c2_mat[c,2,], col=C2_color[c], lwd=1, lty=3)
}

plot_legend()
```


```{r plots_set_5, echo=FALSE}
# > names(cas2_mpd[[1]])
#  [1] "Init"                 "state1"               "estimated_values"
#  [4] "estimated_summary"    "obj_fun"              "Recruitment"
#  [7] "Mortality"            "DerivedQuantities"    "nuisance_Q"
# [10] "SSB"                  "maturity_ogive"       "growth_length_at_age"
# [13] "growth_weight_at_age" "fishery_selectivity"  "CPUE_index"
# [16] "fishery_LF"           "Hess"                 "Covar"
# [19] "Corr"

# plot fishing pressures

max_val <- max(c1_quant$fishing_pressures$pressure,
               c1_sens1_quant$fishing_pressures$pressure,
               max(unlist(list.map(cas2_mpd, max(Mortality$`fishing_pressure[AndesFishery]`)))))

plot(c1_quant$fishing_pressures$year, c1_quant$fishing_pressures$pressure, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='U', main='Fishing pressure comparison: Andes')
lines(c1_sens1_quant$fishing_pressures$year, c1_sens1_quant$fishing_pressures$pressure, type='l', col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Mortality$year, cas2_mpd[[c]]$Mortality$`fishing_pressure[AndesFishery]`, col=C2_color[c], lwd=1)
}

plot_legend()
```


```{r plots_set_6, echo=FALSE}
ages <- seq(1, 80)

max_val <- max(1,
               c1_quant$`Ogive parameter values`$`selectivity[SELandes].all`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[SELandes].all`)
plot(ages, c1_quant$`Ogive parameter values`$`selectivity[SELandes].all`, type='l', col='black', lwd=5, ylim=c(0,max_val), xlab='Age', ylab='', main='Andes fishery selectivity-at-age comparison')

lines(ages, c1_sens1_quant$`Ogive parameter values`$`selectivity[SELandes].all`, col='grey', lwd=2.5)

for (c in 1:num_C2_models)
{
    lines(ages, cas2_mpd[[c]]$fishery_selectivity$Values, col=C2_color[c], lwd=1)
}

plot_legend()
```

