---
title: "Casal2 Test Case comparisons for ORH"
output:
    html_document:
        toc: TRUE
        toc_depth: 2
        number_sections: TRUE

header-includes:
 - \usepackage{pdflscape}
 - \newcommand{\blandscape}{\begin{landscape}}
 - \newcommand{\elandscape}{\end{landscape}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ORH comparison of CASAL and Casal2 model configurations

This document compares the results of at least 2 CASAL model configurations (base and at least one sensitivity) and up to 8 Casal2 model configurations (3 BetaDiff, 2 CppAD, and 3 ADOL-C).

The CASAL model sensitivity 1 has a smaller tolerance value than the CASAL base model (1e-6 vs. 2e-3).

The Casal2 ADOL-C and BetaDiff low tolerance models have a smaller tolerance value than the CASAL base model (1e-6 vs. 2e-3). The Casal2 CppAD models have a tolerance value of 1e-9.


## ORH model characteristics

The main characteristics of the Test Case ORH (orange roughy) CASAL model are:

* one stock, ages 1 - 80
* one area
* years 1979 - 2005, projection years 2006 - 2010
* one time step
* two stock categories: immature and mature, with logistic-producing maturity ogive
* natural mortality (M) of 0.045
* length-weight relationship ($W = aL^b$)
* no stock-recruitment relationship, with $sigma_R$ 1.1
* no ageing error
* one fishery with logistic selectivity-at-age

Observation data include:

* empirical length-at-age data for one year
* one CPUE index
* fishery proportions-at-age

Parameters estimated include:

* B0
* the first and last CV for the age-length relationship
* fishery selectivity parameters a50 and ato95


## R environment

```{r C1_C2_setup, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_set_up_R_env.R')
```


## CASAL and Casal2 model output

```{r CASAL, warning=FALSE}
source('../../R-functions/report_read_in_CASAL_MPD_files.R')
```


```{r Casal2}
source('../../R-functions/report_read_in_Casal2_MPD_files.R')
```



\newpage
\blandscape
## Tables

Tables of parameter estimates and objective function components for the CASAL and Casal2 model MPD results

```{r table_setup, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_set_up_tables.R')

options(scipen=999)
```


```{r huxtables_parameters, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_print_parameter_tables.R')

hux_C1_est_params
hux_C2_est_params
hux_C2_pd_est_params
```


```{r huxtables_obj_fun, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_print_obj_fun_tables.R')

hux_C1_obj_fun
hux_C2_obj_fun
```


```{r convergence_info, echo=FALSE}
source('../../R-functions/report_print_convergence_info.R')
```


```{r warnings_and_bounds, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_print_warnings.R')
```


```{r huxtables_close, echo=FALSE}
options(scipen=0)
```
\elandscape
\newpage



## Matching of outputs

Time series comparisons with CASAL base model results

```{r Time_Series_Match, echo=FALSE, warning=FALSE, message=FALSE}
# do the catch time series differ?

source('../../R-functions/report_time_series_match_function.R')


c1_catch <- c1_quant$actual_catches$catch

for (c in 1:num_C2_models)
{
    print(paste('Catch time series base model comparison for run', C2_subdir[c]))

    print(paste('Actual catches for Andes match:', time_series_match(c1_catch, cas2_mpd[[c]]$Mortality$`catch[AndesFishery]`)))

    print('')
}
```

Derived quantities

SB0, SBcurrent, MSY, F_MSY, others...



## Plots

Comparison plots

```{r plots_base, echo=FALSE}

# c(bottom, left, top, right)
par(mar=c(4,4,2,1) + 0.1)

par(mfrow=c(1,2))

bl_lwd <- 6
gr_lwd <- 3
```


```{r plots_correlation, echo=FALSE}
source('../../R-functions/report_plot_correlation_matrices.R')
```


```{r plots_set_0, echo=FALSE}
source('../../R-functions/report_set_up_legend.R')


max_val <- max(c1_quant$SSBs$SSB,
               c1_sens1_quant$SSBs$SSB,
               max(unlist(list.map(cas2_mpd, max(DerivedQuantities$SSB$values)))))

plot(c1_quant$SSBs$year, c1_quant$SSBs$SSB, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='Biomass (t)', main='SSB comparison')
lines(c1_sens1_quant$SSBs$year, c1_sens1_quant$SSBs$SSB, type='l', col='grey', lwd=gr_lwd)

for (c in 1:num_C2_models)
{
    lines(names(cas2_mpd[[c]]$DerivedQuantities$SSB$values), cas2_mpd[[c]]$DerivedQuantities$SSB$values, col=C2_color[c], lwd=1)
}

plot_legend()
```


```{r plots_set_1, echo=FALSE}

# Q:  how to get CASAL initial age structure?
# A:  use 'casal::extract.partition()''


ages <- seq(1, 80)

# omit the 'category' column
c2_mat <- array(0, dim=c(num_C2_models, dim(as.matrix(cas2_mpd[[1]]$Init$values[,-1]))))
for (c in 1:num_C2_models)
{
    c2_mat[c,,] <- as.matrix(cas2_mpd[[c]]$Init$values[,-1])
}

# plot initial numbers

max_val <- max(cas_part$initial,
               cas_sens1_part$initial,
               max(c2_mat))

plot(ages, cas_part$initial[1,], type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Age', ylab='', main='Initial numbers-at-age comparison (immature and mature)', sub='Dotted lines are for mature')
lines(ages, cas_part$initial[2,], col='black', lwd=bl_lwd, lty=3)

lines(ages, cas_sens1_part$initial[1,], col='grey', lwd=gr_lwd)
lines(ages, cas_sens1_part$initial[2,], col='grey', lwd=gr_lwd, lty=3)

for (c in 1:num_C2_models)
{
    lines(ages, c2_mat[c,1,], col=C2_color[c], lwd=1)
    lines(ages, c2_mat[c,2,], col=C2_color[c], lwd=1, lty=3)
}

plot_legend()



c2_mat <- array(0, dim=c(num_C2_models, dim(as.matrix(cas2_mpd[[1]]$state1$`2005`$values))))
for (c in 1:num_C2_models)
{
    c2_mat[c,,] <- as.matrix(cas2_mpd[[c]]$state1$`2005`$values)
}

# plot final numbers

max_val <- max(cas_part$final,
               cas_sens1_part$final,
               max(c2_mat))

plot(ages, cas_part$final[1,], type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Age', ylab='', main='Final numbers-at-age comparison (immature and mature)', sub='Dotted lines are for mature')
lines(ages, cas_part$final[2,], col='black', lwd=bl_lwd, lty=3)

lines(ages, cas_sens1_part$final[1,], col='grey', lwd=gr_lwd)
lines(ages, cas_sens1_part$final[2,], col='grey', lwd=gr_lwd, lty=3)

for (c in 1:num_C2_models)
{
    lines(ages, c2_mat[c,1,], col=C2_color[c], lwd=1)
    lines(ages, c2_mat[c,2,], col=C2_color[c], lwd=1, lty=3)
}

plot_legend()
```


```{r plots_set_5, echo=FALSE}
# > names(cas2_mpd[[1]])
#  [1] "Init"                 "state1"               "estimated_values"
#  [4] "estimated_summary"    "obj_fun"              "Recruitment"
#  [7] "Mortality"            "DerivedQuantities"    "nuisance_Q"
# [10] "SSB"                  "maturity_ogive"       "growth_length_at_age"
# [13] "growth_weight_at_age" "fishery_selectivity"  "CPUE_index"
# [16] "fishery_LF"           "Hess"                 "Covar"
# [19] "Corr"

# plot fishing pressures

max_val <- max(c1_quant$fishing_pressures$pressure,
               c1_sens1_quant$fishing_pressures$pressure,
               max(unlist(list.map(cas2_mpd, max(Mortality$`fishing_pressure[AndesFishery]`)))))

plot(c1_quant$fishing_pressures$year, c1_quant$fishing_pressures$pressure, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='U', main='Fishing pressure comparison: Andes')
lines(c1_sens1_quant$fishing_pressures$year, c1_sens1_quant$fishing_pressures$pressure, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Mortality$year, cas2_mpd[[c]]$Mortality$`fishing_pressure[AndesFishery]`, col=C2_color[c], lwd=1)
}

plot_legend()
```


```{r plots_set_6, echo=FALSE}
ages <- seq(1, 80)

max_val <- max(1,
               c1_quant$`Ogive parameter values`$`selectivity[SELandes].all`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[SELandes].all`)
plot(ages, c1_quant$`Ogive parameter values`$`selectivity[SELandes].all`, type='l', col='black', lwd=bl_lwd, ylim=c(0,max_val), xlab='Age', ylab='', main='Andes fishery selectivity-at-age comparison')

lines(ages, c1_sens1_quant$`Ogive parameter values`$`selectivity[SELandes].all`, col='grey', lwd=gr_lwd)

for (c in 1:num_C2_models)
{
    lines(ages, cas2_mpd[[c]]$fishery_selectivity$Values, col=C2_color[c], lwd=1)
}

plot_legend()
```

