\section{Running \CNAME\label{sec:RunningCasal2}\index{Running \CNAME}}

\CNAME\ is run from a console window (i.e., the command line) on \I{Microsoft Windows} or from a terminal window on \I{Linux}. \CNAME\ uses information from input data files -{}- the \emph{\config\index{Input configuration file}}\ being the input file that is supplied to \CNAME.

The \config\ is required and defines the model structure, processes, observations, parameters (both the fixed parameters and the parameters to be estimated)\index{Estimable parameters}, and the requested reports (outputs).

By convention, the name of the \config\ ends with the suffix \texttt{.csl2}. However, any suffix is acceptable. The default name for the \config\ is \emph{config.csl2} and if used it does not have to be specified as one of the command line arguments to to \CNAME. Note that the \config\ can include other separate files so the specification can be split into digestible parts.

Command line arguments\index{Command line arguments} are used to specify the actions or \emph{tasks}\index{Tasks} of \CNAME, e.g., to run a model with a set of parameter values, to estimate parameter values (either point estimates or MCMC), to project quantities, or to simulate observations.For example, \texttt{-r} is the \emph{run} mode, \texttt{-e} is the \emph{estimation} mode, and \texttt{-m} is the \emph{MCMC} mode. The \emph{command line arguments} are described in Section \ref{sec:CommandLineArguments}.

\subsection{\I{Using \CNAME}}

To use \CNAME, open a console window (i.e. the command prompt) window on Microsoft Windows or a terminal window on Linux. Navigate to the directory where the model \config s are located. Then enter \CNAME\ with arguments for a specific mode to start the \CNAME\ mode running; see Section \ref{sec:CommandLineArguments} for the list of possible arguments. \CNAME\ will print output to the screen.

For both 64-bit Linux and Microsoft Windows, we recommend using the install package available on \github\ for Microsoft Windows or the Debian package (.deb) for Linux. Running \CNAME\ on a system requires the main binary (casal2.exe on Windows, or casal2 on Linux) and the associated dynamically linked libraries (DLL) for Windows or shared objects (.so) for Linux to be installed into appropriate directories, and cannot be (easily) run by copying the binary to a working directory. \CNAME\ is not available for 32-bit operating systems or macOS.

\subsection{\I{Redirecting standard output}\label{sec:RedirectingStandardOut}}

\CNAME\ uses the \texttt{standard output}\index{standard output} stream to display runtime information. The \I{standard error} stream is used by \CNAME\ to output the program exit status and runtime errors. We suggest redirecting both the standard output and standard error into an appropriate file or files\index{Redirecting standard out}\index{Redirecting standard error}.

With the bash shell (on Linux systems), you can do this using the command structure

\begin{verbatim} (casal2 [arguments] > run.out) >& run.err &\end{verbatim}

It may be useful to redirect the standard input, especially if you're using \CNAME\ inside a batch job, i.e.

\begin{verbatim} (casal2 [arguments] > run.out < /dev/null) >& run.err &\end{verbatim}

On Microsoft Windows systems, you can redirect to standard output using

\begin{verbatim} casal2 [arguments] > run.out\end{verbatim}

And, on some Microsoft Windows systems (e.g., Windows 10), you can redirect to both standard output and standard error, using the syntax

\begin{verbatim} casal2 [arguments] > run.out 2> run.err\end{verbatim}

\CNAME\ outputs header information to the output. The header\index{Output header information} consists of the program name and version, the arguments passed to \CNAME\ from the command line, the date and time that the program was called (derived from the system time), the user name, and the machine name (including the operating system and the process identification number). This information can be used to track output across runs, dates, and versions of \CNAME .

\vspace*{4mm}

\subsection{\I{Command line arguments}\label{sec:CommandLineArguments}}

\CNAME\ is called using:

\texttt{\cname [-c \emph{config\_file}] [\emph{task}] [\emph{options}]}

where

\begin{description}
  \item [\texttt{-c \emph{config\_file}}] Define the \config for \CNAME\ (if this argument is omitted, the default \config\ is \texttt{config.csl2})
\end{description}

and where \emph{task} must be one of the following (\textbf{[]} indicates a secondary label to call the task, e.g. \textbf{\texttt{-h}} will execute the same task as \textbf{\texttt{-{}-help}}),

\begin{description}
\item [\texttt{-h [-{}-help]}] Display help (this page)
\item [\texttt{-l [-{}-licence]}] Display the reference for the software license (GPL v2)
\item [\texttt{-v [-{}-version]}] Display the \CNAME\ version number

\item [\texttt{-r [-{}-run]}] \emph{Run} the model once using the parameter values in the \config, or optionally with the free parameter values from the file specified with argument \texttt{-i} or \texttt{-I}.
\item [\texttt{-e [-{}-estimate]}] Do a point \emph{estimate} using the values in the \config\ as the starting point for the parameters to be estimated, or optionally with the free parameter values from the file specified with argument \texttt{-i} or \texttt{-I}.
\item [\texttt{-E [-{}-Estimate]} \emph{filename}] Do a point \emph{estimate} and generate an MPD file (ie.e., a file containing the free parameters and the covariance matrix). As with \texttt{-e}, this uses the values in the \config\ as the starting point for the parameters to be estimated, or optionally the free parameter values from the file specified with argument \texttt{-i} or \texttt{-I}.
\item [\texttt{-{p}[-profiling]}] Do a likelihood \emph{profile} using the parameter values in the \config\ as the starting point, or optionally with the free parameter values from the file specified with argument \texttt{-i} or \texttt{-I}
\item [\texttt{-m [-{}-mcmc]}] Do an \emph{MCMC}. An estimate run is fist carried out to estimate the co-variance matrix for the MCMC proposal distribution, using the values in the \config\ as the starting point for the parameters to be estimated. Optionally the free parameter values from the file specified with argument \texttt{-i} or \texttt{-I} can be used as the starting point.
\item [\texttt{-M [-{}-mcmc-from-estimate] \emph{filename}}] Do an \emph{MCMC} using the covariance and free parameters in the MPD file.
\item [\texttt{-R [-{}-resume] \emph{filename}}] Resume a previously stopped\emph{MCMC} using the covariance and free parameters in the MPD file. Additional arguments must be supplied to specify the sample and objective files from the previous MCMC with \texttt{-{}-objective-file} and \texttt{-{}-sample-file}.
\item [\texttt{-f [-{}-projection]}] \emph{n}. Project the model \emph{forward} in time using the parameter values in the \config\ as the starting point for the estimation, or optionally with the parameter values from the file specified with the argument \texttt{-i} or \texttt{-I}. Projections are repeated for each parameter set (i.e., each line of data in the free parameter file) \emph{n} times (the default is 1). Typically, the MCMC sample output will be used with \texttt{-i}.
\item [\texttt{-s [-{}-simulation]} \emph{n}] \emph{Simulate} \emph{n} of observation sets using values in the \config\ as the parameter values, or optionally with the parameter values from the file specified with the argument \texttt{-i} or \texttt{-I}.

\end{description}

The following optional arguments\index{Optional command line arguments} [\emph{options}] may be specified

\begin{description}
\item [\texttt{-i [-{}-input] \emph{filename}}] \emph{Input} one or more sets of free (estimated) parameter values from \texttt{\emph{filename}} (see Section \ref{sec:Report} for details about the format of \texttt{\emph{filename}}).
\item [\texttt{-I [-{}-input-force] \emph{filename}}] \emph{Input} one or more sets of parameter values from \texttt{\emph{filename}}. This contains both the free parameters and also force the \emph{overwrite} addressable (non-estimated) values in the \config\ (see Section \ref{sec:Report} for details about the format of \texttt{\emph{filename}})
\item [\texttt{-o [-{}-output] \emph{filename}}] \emph{Output} a report of the free (estimated) parameter values in a format suitable for \texttt{-i \emph{filename}} (see Section \ref{sec:Report} for details about the format of \texttt{\emph{filename}})
\item [\texttt{-g [-{}-seed] \emph{seed}}] Initialise the random number \emph{generator} with \texttt{\emph{seed}}, a positive (long) integer value (note, if \texttt{-g} is not specified, then \CNAME\ will  generate a random number seed based on the computer clock time)
\item [\texttt{-{}-loglevel \emph{arg}}] Set the level for information or logging messages from \CNAME. Valid options are (from more verbose to less verbose) trace, finest, fine, medium, info, important, and warning. The default is 'info' (see Section \ref{sec:Report} for more information).
\item [\texttt{-t [-{}-tabular]}] Print \command{report} in tabular format (see Section \ref{sec:Report} for more information).
\item [\texttt{-{}-single-step}] Run with \texttt{-r} to pause the model and ask the user to specify parameters and their values to use for the next iteration (see Section \ref{sec:SingleStepping})
\item [\texttt{-q [-{}-query]\emph{object type}}] \emph{Query} an object type to print an extract of the object description and parameter definitions.  An object can be defined as \texttt{\emph{block.type}}, e.g., \texttt{casal2 -{}-query process.recruitment\_constant} will query the constant recruitment block.

\end{description}
%
Combinations of these command line arguments can also be implemented. Examples of some useful ones are below.
\begin{description}
	\item \texttt{casal -r -i par.file> mulit\_run.out} will conduct multiple model runs, one for each row of parameters in \texttt{par.file}. This can be useful for investigating the effect of individual parameters in the model or summarising profiled outputs.
	
	\item \texttt{casal -e -i par.file> mulit\_estimate.out} will conduct multiple estimation routines. One for each row of parameters in \texttt{par.file}. This can be useful for assessing convergence to a global minimum. All base models should be run from multiple starting parameter values to assess model convergence/sensitivity to starting values.
	
	\item \texttt{casal -s 10 -i par.file > mulit\_simulation.out} This command instructs \CNAME\ to simulate 10 sets of simulated data sets for each each row of parameters in \texttt{par.file}. The \texttt{-s} component adds observation error in simulated data sets through the likelihood distribution assumptions and the \texttt{-i} adds parameter uncertainty into the simulated data sets if each row differs.
\end{description}


\subsection{Constructing the \CNAME\ \config s \label{ConstructingConfig}}\index{Input configuration file syntax}

\begin{itemize}
	\item the description of the population structure, dynamics, and parameters. See Section \ref{sec:Population},
	\item the estimation methods and estimated variables. See Section \ref{sec:Estimation},
	\item the observations and their associated properties and likelihoods. See Section \ref{sec:Observation}, and
	\item the results that \CNAME\ will output. See Section \ref{sec:Report}.
\end{itemize}

Note that \config\ files can \emph{include} other \config\ files to assist file management, using the command \texttt{!include "\emph{filename}"}. See Section \ref{syntax:General} for more details.

\subsubsection{Commands}\index{Commands}

\CNAME\ has a range of commands that define the model structure, processes, parameters, observations, and how tasks are carried out. There are three types of commands

\begin{itemize}
	\item Commands that have an argument only and do not have subcommands (for example, \texttt{!include}\ \argument{\emph{filename}})
	\item Commands that have a label and subcommands (for example \command{process} must have a label and has subcommands)
	\item Commands that do not have either a label or argument, but have subcommands (for example \command{model} or \command{categories})
\end{itemize}

Apart from \texttt{!include}, commands start with an \texttt{@} in the first column (i.e., may not have a space or tab character before them on the line). After each command, the subcommands are listed and must occur before the next command. Otherwise, the commands and subcommands are free form with each command or subcommand on a separate line (see Section \ref{sec:CommandBlockFormat}).

Commands that have a label must have a unique label, i.e., the label cannot be used on more than one command of that type. \CNAME\ checks and will report an error of two commands of the same type have the same label. The labels can contain alpha numeric characters, period (`.'), underscore (`\_') and dash (`-'), but cannot start with a double underscore ('\_\_'). Labels that start with a double underscore are reserved, and used for internal reports that \CNAME\ can automatically generate in some circumstances. Otherwise labels must not contain white space (tabs or spaces) or any characters that are not letters, numbers, dashes, periods, or underscores. For example,

{\small{\begin{verbatim}
@process NaturalMortality
\end{verbatim}}}
or
{\small{\begin{verbatim}
!include MyModelSpecification.csl2
\end{verbatim}}}

\subsubsection{Subcommands}\index{Commands ! Subcommands}

\CNAME\ subcommands define options and parameter values related to a particular command. Subcommands always take an argument which is one of a specific \emph{type}. The \emph{types} for each subcommand are defined in Section \ref{sec:syntax}, and are summarised below.

Like commands (\command{command}), subcommands and their arguments are not order specific, except that that all subcommands of a given command must appear before the next \command{command} block. \CNAME\ may report an error if they are not supplied in this way. However, in some circumstances a different order may result in a valid, but unintended, set of actions, leading to unexpected results.

The argument type for a subcommand can be\index{Subcommand argument type}:

\begin{description}
    \item[switch] true/false
    \item[integer] an integer number
	\item[integer vector] a vector of integer numbers
	\item[integer range] a range of integer numbers separated by a colon, e.g. 1994:1996 is expanded to an integer vector of values (1994 1995 1996)
	\item[constant] a real number (i.e., a double)
	\item[constant vector] a vector of real numbers (i.e., a vector of doubles)
	\item[estimable] a real number that can be estimated (i.e., a double)
	\item[estimable vector] a vector of real numbers that can be estimated (i.e., a vector of doubles)
	\item[addressable] a real number that can be referenced but not estimated (i.e., an addressable double)
	\item[addressable vector] a vector of real numbers that can be referenced but not estimated (i.e., a vector of addressable doubles)
	\item[string] a categorical (string) value
	\item[string vector] a vector of categorical values
\end{description}

Switches are characteristics which are either true or false. Enter \emph{true} as \argument{true} or \argument{t}, and \emph{false} as \argument{false} or \argument{f}.

Integers must be entered as whole numbers without decimal points (i.e., if \subcommand{year}\ is an integer then it is specified as \texttt{2008}, not \texttt{2008.0})

Arguments of type integer vector, constant vector, estimable vector, addressable vector, or categorical vector must contain one or more entries on a row, separated by white space (tabs or spaces). Arguments of type integer range must contain a colon (\texttt{:}) and no white space (tabs or spaces).

Parameters are defined in the population section and most (but not all) numeric parameters can be estimated. See Section \ref{sec:syntax} for the list of available parameters and if they are allowed to be estimated. Note that parameters will only be estimated if requested using an \command{estimate} command, and are otherwise treated as a constant.

Parameters can also be addressable, i.e., they can be referred to within another command or command block by using their addressable name. See Section \ref{sec:syntax} to determine of a subcommand is addressable.

\subsubsection{The command block format}\index{Command block format}\label{sec:CommandBlockFormat}

The command block is a basic unit within the \config. Each command begins with the symbol \command{} and then the command name, usually followed by a user defined label or a valid argument. The end of each command block is denoted by the start of the next command block or end of the file. For example, the layout of a \config\ will be

\begin{description}
	\item \command{command} \subcommand{label}
	\item \subcommand{first\_subcommand} \subcommand{argument}
	\item \subcommand{second\_subcommand} \subcommand{argument}
	\item ... etc.
	\item \command{another\_command} \subcommand{label}
	\item \subcommand{another\_subcommand} \subcommand{argument}
	\item \subcommand{another\_subcommand} \subcommand{argument}
	\item ... etc.
\end{description}

Note that subcommands can be in any order within each command block. And command blocks can be in any order within the input files, except \command{model} --- this must be the first command block encountered by \CNAME.

Blank lines are ignored, as is extra white space (tabs and spaces) between arguments. However, to start command block the \command{} character must be the first character on the line and must not be preceded by any white space. Each input file must end with a carriage return.

Commands, subcommands, and arguments in the \config s are not case sensitive. However, labels and variable values are case sensitive. Note that on Linux (unlike Microsoft Windows) specification of any file names or file paths will be case sensitive.

\subsubsection{\I{Commenting out lines}}\index{Comments}

Text on a line that starts with the symbol \commentline\ is considered to be a comment and is ignored. To comment out a group of commands or subcommands, use \commentline\ at the beginning of each line to be ignored.

Alternatively, to comment out an entire block or section, use \commentstart\ at the beginning of a line to start the comment block, then end the block with \commentend. All lines (including line breaks) between \commentstart\ and \commentend\ inclusive are ignored.

\small{\begin{verbatim}
  # This line is a comment and will be ignored
  @process NaturalMortality
  m 0.2
  /* This block of text
  is a comment and
  will be ignored
  */
\end{verbatim}}

\subsubsection{How to reference parameters\label{sec:parameter-names}\index{Determining parameter names}\index{Parameter names}}

All parameters have an unique name, allowing it to be referenced in other command blocks. When \CNAME\ processes the \config\ it translates each command block (see section~\ref{sec:CommandBlockFormat}) and each subcommand block into an object, each with a unique parameter name. For commands, this parameter name is simply the command label. For subcommands, the parameter name format is one of the following:

\begin{description}
	\item \texttt{command[label].subcommand} if the command has a label, or
	\item \texttt{command.subcommand} if the command has no label, or
	\item \texttt{command[label].subcommand\{i\}} if the command has a label and the subcommand arguments are a vector, and we are accessing the  \emph{i}th element of that vector.
	\item \texttt{command[label].subcommand\{i:j\}} if the command has a label, and the subcommand arguments are a vector, and we are accessing the elements from $i$ to $j$ (inclusive) of that vector.
\end{description}

For example, the parameter name of a process of instantaneous mortality (i.e., natural mortality) is the subcommand \subcommand{m} of a \command{process} of type \subcommand{mortality\_constant\_rate}, i.e., the command block may be

\small{\begin{verbatim}
				@process NaturalMortality
				type mortality_constant_rate
				categories male female
				m 0.2 0.2
\end{verbatim}}
%
\texttt{process[NaturalMortality].m} is the unique reference for the vector of male then females natural mortality values (\textbf{note:} order will follow categories order). But to reference just the 'female' rate then the form is
\texttt{process[NaturalMortality].m\{female\}}.

\subsection{\I{Reading a command block}\label{sec:Readingcommandblock}}\index{Reading\ command\ block}\index{Reading\ command\ block section}

Here, we illustrate reading a command block using two important commands, \texttt{@process} and \texttt{@estimate}.

The command \texttt{@process} specifies a process that can be used in the model. The are a fixed set of predefined processes (subroutines in C++ code). The way to identify which one is by the \texttt{type} subcommand. Processes can take one or more parameters and some will need other data to be applied too. Some parameters are mandatory and others can take a default value if they are not specified.
We have categories male and female, and two fisheries, line and pot. The command block starts with a\textit{@process}:

{\small{\begin{verbatim}
@process Fishing
type mortality_instantaneous
\end{verbatim}}}

This sets up a process block using the \textit{mortality\_instantaneous} process which simultaneously depletes the population by natural mortality and from two types of fishing. Its label is \textit{Fishing}.

Next we specify the values for natural mortality (\textit{m}) an argument to this process, to 0.17 and specify that fisheries acts on all categories. Note there are two values for natural mortality, one for each sex. The parameters \textit{m} can be estimated, if required. The command block fragment:

{\small{\begin{verbatim}
m 0.17 0.17   # natural mortality  for each category
relative_m_by_age One One # natural mortality multiplier
categories *  # fishing acts on all categories ("*" shorthand for male female)
\end{verbatim}}}

Catches are supplied via a \textit{table} format using three columns: year and two for the fisheries, one for each, which take the labels \textit{line} and \textit{pot}. Column names are on the first line of the table and these columns can be in any order,

{\small{\begin{verbatim}
#catches
table catch         # define catches by fishery in table format
year line pot       #names columns so can identity catch for each fishery
2000 1000 2000      # catches by year
2001 500  1000
2002 1000 5000
end_table           # end of table marker
\end{verbatim}}}

Other information needed are supplied in the methods table which has a fixed number of columns (again these can be in any order), one for each piece of information needed to model a fishery. The method column defines the fishery name which is used in the catch table and also in other observations like age composition from that fishery. The categories that the fishery operates on (all in this case, but it could be just males for one and females for the other) are in the category column, the fishing selectivity to be used is given as a selectivity block name which is define somewhere else in the files, $U_max$ is the maximum exploitation rate in any year that is allowed, next comes the time step the fishing operates in, and lastly the block name of a penalty function that is used to penalise estimable parameter values that result in the supplied catch not being caught. Again, the penalty block is define elsewhere in the files. After the row with the column names, there follows one row for each fishery:

{\small{\begin{verbatim}
table method        # supply arguments and name selectivity etc
method     category  selectivity    u_max    time_step      penalty
pot        *         potFSel        0.7      1              CatchMustBeTaken1
line       *         lineFSel       0.7      1              CatchMustBeTaken1
end_table
\end{verbatim}}}

To estimate natural mortality, you need to supply an \textit{@estimate} block with a reference name back to \textit{m} in the \textit{Fishing} block. For \textit{@estimate}, \textit{type} specifies the prior to be used in the estimation, which in this case is a normal distribution:

{\small{\begin{verbatim}
@estimate estimate.m
type normal # prior type
parameter process[fishing].m  #
    \*Fishing is unique amongst the @process command blocks
      so this defines the unique reference to the parameter m
    *\

mu  0.2 0.2   #argument to prior = mean
sd  0.02 0.02 #another argument to the prior = standard deviation
\end{verbatim}}}

Note that there are two \textit{m}s, one for each sex, so there has to be two priors. The \textit{\@estiamte} label \textit{estimate.m} is often redundant, but it may be needed in some circumstances.

To estimate a common \textit{m} over both sexes, we estimate one \textit{m}, say the female category, and use the \textit{same} subcommand to apply the same value to the male category \textit{m},

{\small{\begin{verbatim}
@estimate estimate.m
type normal
parameter process[fishing].m{male}
# {} is used to index one or more elements in a vector
same process[fishing].m{female} # set female values = male estimated values
# The mean of the prior
mu  0.2
# The standard deviation of the prior
sd  0.02
\end{verbatim}}}

\subsection{\I{Single-stepping \CNAME}\label{sec:SingleStepping}}\index{Single\_stepping}\index{single\_stepping section}


\TODO{Move this section to somewhere else?}

Single-stepping means \CNAME\ can 'pause' after each year in the annual cycle during a model run, write reports, then wait and process user input of updated estimable parameters for the next year (see the command line argument \texttt{ -{}-single-step}).

This enables \CNAME\ to implement models for management simulations or scenarios that require feedback and can be used, for example, in operational management procedures (OMPs). The single-stepping process can be automated using \R, so that \CNAME\ may be used with \R\ to update input harvest values (e.g., catches from a fishery in a fisheries model) to evaluate a particular harvest control rule.

\subsection{\I{Validating models across versions}}\label{sec:Assert}\index{Asserts}\index{Unit tests}

\CNAME\ has a number of built-in capabilities to validate and verify the code across versions. Unit tests of the underlying \CNAME\ code are carried out at build time, using the \href{https://github.com/google/googletest}{Google Test and Mock} unit testing and mocking framework. The unit test framework aims to cover a significant proportion of the key functionality within the \CNAME\ code base. The unit test code\index{Unit tests} for \CNAME\ is available as a part of the underlying source code.

\CNAME\ can also validate or check certain addressables parameters as a part of testing and validation with the assert command. Asserts check the value of a specific addressables (for example, and observations, parameters, or the objective function). Asserts are one aspect of the internal tests \CNAME\ uses to ensure accuracy across versions and revisions (see Section \ref{syntax:Assert})

\subsection{\CNAME\ exit status values\index{Exit status value}}

When \CNAME\ is run, it will either complete its task successfully or output an error. \CNAME\ will return a single exit status value 'completed' to the standard output. Error messages will be printed to the console. When input file configuration errors are found, \CNAME\ will print error messages, along with the associated filename(s) and line number(s) where the errors were identified, for example,

{\small{\begin{verbatim}
	[ERROR] At line 15 in Reports.csl2: Parameter '{' is not supported
\end{verbatim}}}
