---
title: "Casal2 Test Case comparisons for HOK"
output:
    html_document:
        toc: TRUE
        toc_depth: 2
        number_sections: TRUE

header-includes:
 - \usepackage{pdflscape}
 - \newcommand{\blandscape}{\begin{landscape}}
 - \newcommand{\elandscape}{\end{landscape}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# HOK comparison of CASAL and Casal2 model configurations

This document compares the results of at least 2 CASAL model configurations (base and at least one sensitivity) and up to 8 Casal2 model configurations (3 BetaDiff, 2 CppAD, and 3 ADOL-C).

The CASAL model sensitivity 1 has a smaller tolerance value than the CASAL base model (1e-9 vs. 1e-6).

The Casal2 ADOL-C and BetaDiff low tolerance models have a smaller tolerance value than the CASAL base model (1e-9 vs. 1e-6). The Casal2 CppAD models have a tolerance value of 1e-9.


## HOK model characteristics

The main characteristics of the Test Case HOK (hoki) CASAL model are:

* one stock, ages 1 - 17
* one area, although many characteristics are separated into "east" and "west"
* years 1972 - 2018, projection years 2019 - 2023
* five time steps: Oct_Nov, Dec_Mar, Apr_Jun, End_Jun, and Jul_Sep
* five stock categories: west.sa, east.cr, west.cr, west.wc, and east.cs (labeled "stock.area")
* five migration processes, which have age-specific proportions specified
* sex-specific natural mortality-at-age is double exponential
* area-specific von Bertalanffy age-length relationship
* area-specific length-weight relationship ($W = aL^b$)
* area-specific Beverton-Holt stock-recruitment relationships, with steepness (h) set to 0.75
* ageing error is specified
* surveys with double normal selectivity
* six area-specific fisheries with double normal selectivity

Observation data include:

* survey indices
* survey proportions-at-age
* fishery proportions-at-age

Parameters estimated include:

* area-specific B0
* parameters for the natural mortality-at-age ogive
* free catchability (q) for the surveys
* parameters for the selectivity curves for the surveys
* parameters for the selectivity curves for the fisheries
* migration proportions by age
* area-specific year-class strengths (recruitment deviations)


## R environment

```{r C1_C2_setup, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_set_up_R_env.R')
```


## CASAL and Casal2 model output

```{r CASAL, warning=FALSE}
source('../../R-functions/report_read_in_CASAL_MPD_files.R')
```


```{r Casal2}
source('../../R-functions/report_read_in_Casal2_MPD_files.R')
```



\newpage
\blandscape
## Tables

Tables of parameter estimates and objective function components for the CASAL and Casal2 model MPD results

```{r table_setup, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_set_up_tables.R')

options(scipen=999)
```


```{r huxtables_parameters, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_print_parameter_tables.R')

hux_C1_est_params
hux_C2_est_params
hux_C2_pd_est_params
```


```{r huxtables_obj_fun, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_print_obj_fun_tables.R')

hux_C1_obj_fun
hux_C2_obj_fun
```


```{r convergence_info, echo=FALSE}
source('../../R-functions/report_print_convergence_info.R')
```


```{r warnings_and_bounds, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_print_warnings.R')
```


```{r huxtables_close, echo=FALSE}
options(scipen=0)
```
\elandscape
\newpage



## Matching of outputs

Time series comparisons with CASAL base model results

```{r Time_Series_Match, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_time_series_match_function.R')


for (c in 1:num_C2_models)
{
    print(paste('Catch time series base model comparison for run', C2_subdir[c]))

    print(paste('Actual catches for Ensp1 match:',
                time_series_match(c1_quant$actual_catches$Ensp1, cas2_mpd[[c]]$Mortality$`actual_catch[Ensp1]`)))

    print(paste('Actual catches for Wnsp1 match:',
                time_series_match(c1_quant$actual_catches$Wnsp1, cas2_mpd[[c]]$Mortality$`actual_catch[Wnsp1]`)))

    print(paste('Actual catches for Ensp2 match:',
                time_series_match(c1_quant$actual_catches$Ensp2, cas2_mpd[[c]]$Mortality$`actual_catch[Ensp2]`)))

    print(paste('Actual catches for Wnsp2 match:',
                time_series_match(c1_quant$actual_catches$Wnsp2, cas2_mpd[[c]]$Mortality$`actual_catch[Wnsp2]`)))

    print(paste('Actual catches for Esp match:',
                time_series_match(c1_quant$actual_catches$Esp, cas2_mpd[[c]]$Mortality$`actual_catch[Esp]`)))

    print(paste('Actual catches for Wsp match:',
                time_series_match(c1_quant$actual_catches$Wsp, cas2_mpd[[c]]$Mortality$`actual_catch[Wsp]`)))

    print('')
}
```

Derived quantities

SB0, SBcurrent, MSY, F_MSY, others...



## Plots

Comparison plots

```{r plots_base, echo=FALSE}

# c(bottom, left, top, right)
par(mar=c(4,4,2,1) + 0.1)

par(mfrow=c(1,2))
```


```{r plots_correlation, echo=FALSE}
source('../../R-functions/report_plot_correlation_matrices.R')
```


```{r plots_set_0, echo=FALSE}
source('../../R-functions/report_set_up_legend.R')

# plot E SSB

max_val <- max(c1_quant$SSBs$E,
               c1_sens1_quant$SSBs$E,
               max(unlist(list.map(cas2_mpd, max(SSB$SSB_E$values)))))

plot(c1_quant$SSBs$year, c1_quant$SSBs$E, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='Biomass (t)', main='E SSB comparison')
lines(c1_sens1_quant$SSBs$year, c1_sens1_quant$SSBs$E, type='l', col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(names(cas2_mpd[[c]]$SSB$SSB_E$values), cas2_mpd[[c]]$SSB$SSB_E$values, col=C2_color[c], lwd=1)
}

plot_legend()


# plot W SSB

max_val <- max(c1_quant$SSBs$W,
               c1_sens1_quant$SSBs$W,
               max(unlist(list.map(cas2_mpd, max(SSB$SSB_W$values)))))

plot(c1_quant$SSBs$year, c1_quant$SSBs$W, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='Biomass (t)', main='W SSB comparison')
lines(c1_sens1_quant$SSBs$year, c1_sens1_quant$SSBs$W, type='l', col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(names(cas2_mpd[[c]]$SSB$SSB_E$values), cas2_mpd[[c]]$SSB$SSB_W$values, col=C2_color[c], lwd=1)
}

plot_legend()
```


```{r plots_set_1, echo=FALSE}
# plot initial numbers

# Q:  how to get CASAL initial age structure?
# A:  use 'casal::extract.partition()''



# omit the 'category' column
# categories: west.sa, east.cr, west.cr, west.wc, east.cs

c2_mat <- array(0, dim=c(num_C2_models, dim(as.matrix(cas2_mpd[[1]]$Init$values[,-1]))))
for (c in 1:num_C2_models)
{
    c2_mat[c,,] <- as.matrix(cas2_mpd[[c]]$Init$values[,-1])
}

for (i in 1:nrow(c2_mat[1,,]))
{
    max_val <- max(cas_part$initial[i,],
                   cas_sens1_part$initial[i,],
                   max(c2_mat[,i,]))

    plot(seq(1, 17), cas_part$initial[i,], type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Age', ylab='', main=paste('Initial numbers-at-age comparison:', cas2_mpd[[1]]$Init$values[i,1]))
    lines(seq(1, 17), cas_sens1_part$initial[i,], col='grey', lwd=2.5)
    for (c in 1:num_C2_models)
    {
        lines(seq(1, dim(c2_mat[c,,])[2]), c2_mat[c,i,], col=C2_color[c], lwd=1)
    }

    plot_legend()
}



c2_mat <- array(0, dim=c(num_C2_models, dim(as.matrix(cas2_mpd[[1]]$state5$`2018`$values))))
for (c in 1:num_C2_models)
{
    c2_mat[c,,] <- as.matrix(cas2_mpd[[c]]$state5$`2018`$values)
}

# plot final year numbers (time step 5)

for (i in 1:nrow(c2_mat[1,,]))
{
    max_val <- max(cas_part$final[i,],
                   cas_sens1_part$final[i,],
                   max(c2_mat[,i,]))

    plot(seq(1, 17), cas_part$final[i,], type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Age', ylab='', main=paste('Final numbers-at-age comparison:', cas2_mpd[[1]]$Init$values[i,1]))
    lines(seq(1, 17), cas_sens1_part$final[i,], col='grey', lwd=2.5)
    for (c in 1:num_C2_models)
    {
        lines(seq(1, dim(c2_mat[c,,])[2]), c2_mat[c,i,], col=C2_color[c], lwd=1)
    }

    plot_legend()
}
```


```{r plots_set_2, echo=FALSE}
# plot YCS

max_val <- max(c1_quant$YCS$E,
               c1_sens1_quant$YCS$E,
               max(unlist(list.map(cas2_mpd, max(Recruit_E$ycs_values)))))

plot(c1_quant$YCS$E, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='', main='E YCS comparison')
lines(c1_sens1_quant$YCS$E, type='l', col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Recruit_E$ycs_values, col=C2_color[c], lwd=1)
}

plot_legend()


max_val <- max(c1_quant$YCS$W,
               c1_sens1_quant$YCS$W,
               max(unlist(list.map(cas2_mpd, max(Recruit_W$ycs_values)))))

plot(c1_quant$YCS$W, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='', main='W YCS comparison')
lines(c1_sens1_quant$YCS$W, type='l', col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Recruit_W$ycs_values, col=C2_color[c], lwd=1)
}

plot_legend()



# plot true YCS

max_val <- max(c1_quant$true_YCS$E,
               c1_sens1_quant$true_YCS$E,
               max(unlist(list.map(cas2_mpd, max(Recruit_E$true_ycs)))))

plot(c1_quant$true_YCS$year, c1_quant$true_YCS$E, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='', main='E true YCS comparison')
lines(c1_sens1_quant$true_YCS$year, c1_sens1_quant$true_YCS$E, type='l', col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Recruit_E$ycs_years, cas2_mpd[[c]]$Recruit_E$true_ycs, col=C2_color[c], lwd=1)
}

plot_legend()


max_val <- max(c1_quant$true_YCS$W,
               c1_sens1_quant$true_YCS$W,
               max(unlist(list.map(cas2_mpd, max(Recruit_W$true_ycs)))))

plot(c1_quant$true_YCS$year, c1_quant$true_YCS$W, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='', main='W true YCS comparison')
lines(c1_sens1_quant$true_YCS$year, c1_sens1_quant$true_YCS$W, type='l', col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Recruit_E$ycs_years, cas2_mpd[[c]]$Recruit_W$true_ycs, col=C2_color[c], lwd=1)
}

plot_legend()
```


```{r plots_set_3, echo=FALSE}
# plot recruitment

max_val <- max(c1_quant$recruitments$E,
               c1_sens1_quant$recruitments$E,
               max(unlist(list.map(cas2_mpd, max(Recruit_E$Recruits)))))

plot(c1_quant$recruitments$year, c1_quant$recruitments$E, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='', main='E Recruits comparison')
lines(c1_sens1_quant$recruitments$year, c1_sens1_quant$recruitments$E, type='l', col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Recruit_E$ycs_years, cas2_mpd[[c]]$Recruit_E$Recruits, col=C2_color[c], lwd=1)
}

plot_legend()


max_val <- max(c1_quant$recruitments$W,
               c1_sens1_quant$recruitments$W,
               max(unlist(list.map(cas2_mpd, max(Recruit_W$Recruits)))))

plot(c1_quant$recruitments$year, c1_quant$recruitments$W, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='', main='W Recruits comparison')
lines(c1_sens1_quant$recruitments$year, c1_sens1_quant$recruitments$W, type='l', col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Recruit_E$ycs_years, cas2_mpd[[c]]$Recruit_W$Recruits, col=C2_color[c], lwd=1)
}

plot_legend()
```


```{r plots_set_4, echo=FALSE}
# plot surveys

c1_surv      <- cas_mpd$free$`q[WCacous].q` * unlist(c1_quant$WCacous)
c1_s1_surv   <- cas_mpd_sens1$free$`q[WCacous].q` * unlist(c1_sens1_quant$WCacous)

c2_surv <- list()
for (c in 1:num_C2_models)
{
    c2_surv[[c]] <- cas2_mpd[[c]]$WCacous$Values
}

max_val <- max(c1_surv,
               c1_s1_surv,
               (c2_surv[[1]]$observed * exp(1.96 * c2_surv[[1]]$error_value)),
               max(unlist(list.map(c2_surv, max(expected)))))

plot(names(c1_surv), c1_surv, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='Biomass (t)', main='Survey comparison: WCacous')
lines(names(c1_s1_surv), c1_s1_surv, type='l', col='grey', lwd=2.5)
points(c2_surv[[c]]$year, c2_surv[[c]]$observed, pch=20, col='black')
arrows(c2_surv[[c]]$year, (c2_surv[[c]]$observed * exp(-1.96 * c2_surv[[c]]$error_value)), c2_surv[[c]]$year, (c2_surv[[c]]$observed * exp(1.96 * c2_surv[[c]]$error_value)), length=0.05, angle=90, code=3)
for (c in 1:num_C2_models)
{
    points(c2_surv[[c]]$year, c2_surv[[c]]$expected, col=C2_color[c], pch=(15+c-1))
}

plot_legend()



c1_surv      <- cas_mpd$free$`q[CRsum].q` * unlist(c1_quant$CRsumbio)
c1_s1_surv   <- cas_mpd_sens1$free$`q[CRsum].q` * unlist(c1_sens1_quant$CRsumbio)

c2_surv <- list()
for (c in 1:num_C2_models)
{
    c2_surv[[c]] <- cas2_mpd[[c]]$CRsumbio$Values
}

max_val <- max(c1_surv,
               c1_s1_surv,
               (c2_surv[[1]]$observed * exp(1.96 * c2_surv[[1]]$error_value)),
               max(unlist(list.map(c2_surv, max(expected)))))

plot(names(c1_surv), c1_surv, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='Biomass (t)', main='Survey comparison: CRsumbio')
lines(names(c1_s1_surv), c1_s1_surv, type='l', col='grey', lwd=2.5)
points(c2_surv[[c]]$year, c2_surv[[c]]$observed, pch=20, col='black')
arrows(c2_surv[[c]]$year, (c2_surv[[c]]$observed * exp(-1.96 * c2_surv[[c]]$error_value)), c2_surv[[c]]$year, (c2_surv[[c]]$observed * exp(1.96 * c2_surv[[c]]$error_value)), length=0.05, angle=90, code=3)
for (c in 1:num_C2_models)
{
    points(c2_surv[[c]]$year, c2_surv[[c]]$expected, col=C2_color[c], pch=(15+c-1))
}

plot_legend()



c1_surv      <- cas_mpd$free$`q[SAsum].q` * unlist(c1_quant$SAbio)
c1_s1_surv   <- cas_mpd_sens1$free$`q[SAsum].q` * unlist(c1_sens1_quant$SAbio)

c2_surv <- list()
for (c in 1:num_C2_models)
{
    c2_surv[[c]] <- cas2_mpd[[c]]$SAsumbio$Values
}

max_val <- max(c1_surv,
               c1_s1_surv,
               (c2_surv[[1]]$observed * exp(1.96 * c2_surv[[1]]$error_value)),
               max(unlist(list.map(c2_surv, max(expected)))))

plot(names(c1_surv), c1_surv, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='Biomass (t)', main='Survey comparison: CRsumbio')
lines(names(c1_s1_surv), c1_s1_surv, type='l', col='grey', lwd=2.5)
points(c2_surv[[c]]$year, c2_surv[[c]]$observed, pch=20, col='black')
arrows(c2_surv[[c]]$year, (c2_surv[[c]]$observed * exp(-1.96 * c2_surv[[c]]$error_value)), c2_surv[[c]]$year, (c2_surv[[c]]$observed * exp(1.96 * c2_surv[[c]]$error_value)), length=0.05, angle=90, code=3)
for (c in 1:num_C2_models)
{
    points(c2_surv[[c]]$year, c2_surv[[c]]$expected, col=C2_color[c], pch=(15+c-1))
}

plot_legend()
```


```{r plots_set_5, echo=FALSE}
# plot fishing pressures

max_val <- max(c1_quant$fishing_pressures$Ensp1,
               c1_sens1_quant$fishing_pressures$Ensp1,
               max(unlist(list.map(cas2_mpd, max(Mortality$`fishing_pressure[Ensp1]`)))))

plot(c1_quant$fishing_pressures$year, c1_quant$fishing_pressures$Ensp1, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='U', main='Fishing pressure comparison: Ensp1')
lines(c1_sens1_quant$fishing_pressures$year, c1_sens1_quant$fishing_pressures$Ensp1, type='l', col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Mortality$year, cas2_mpd[[c]]$Mortality$`fishing_pressure[Ensp1]`, col=C2_color[c], lwd=1)
}

plot_legend()



max_val <- max(c1_quant$fishing_pressures$Wnsp1,
               c1_sens1_quant$fishing_pressures$Wnsp1,
               max(unlist(list.map(cas2_mpd, max(Mortality$`fishing_pressure[Wnsp1]`)))))

plot(c1_quant$fishing_pressures$year, c1_quant$fishing_pressures$Wnsp1, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='U', main='Fishing pressure comparison: Wnsp1')
lines(c1_sens1_quant$fishing_pressures$year, c1_sens1_quant$fishing_pressures$Wnsp1, type='l', col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Mortality$year, cas2_mpd[[c]]$Mortality$`fishing_pressure[Wnsp1]`, col=C2_color[c], lwd=1)
}

plot_legend()



max_val <- max(c1_quant$fishing_pressures$Ensp2,
               c1_sens1_quant$fishing_pressures$Ensp2,
               max(unlist(list.map(cas2_mpd, max(Mortality$`fishing_pressure[Ensp2]`)))))

plot(c1_quant$fishing_pressures$year, c1_quant$fishing_pressures$Ensp2, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='U', main='Fishing pressure comparison: Ensp2')
lines(c1_sens1_quant$fishing_pressures$year, c1_sens1_quant$fishing_pressures$Ensp2, type='l', col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Mortality$year, cas2_mpd[[c]]$Mortality$`fishing_pressure[Ensp2]`, col=C2_color[c], lwd=1)
}

plot_legend()



max_val <- max(c1_quant$fishing_pressures$Wnsp2,
               c1_sens1_quant$fishing_pressures$Wnsp2,
               max(unlist(list.map(cas2_mpd, max(Mortality$`fishing_pressure[Wnsp2]`)))))

plot(c1_quant$fishing_pressures$year, c1_quant$fishing_pressures$Wnsp2, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='U', main='Fishing pressure comparison: Wnsp2')
lines(c1_sens1_quant$fishing_pressures$year, c1_sens1_quant$fishing_pressures$Wnsp2, type='l', col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Mortality$year, cas2_mpd[[c]]$Mortality$`fishing_pressure[Wnsp2]`, col=C2_color[c], lwd=1)
}

plot_legend()



max_val <- max(c1_quant$fishing_pressures$Esp,
               c1_sens1_quant$fishing_pressures$Esp,
               max(unlist(list.map(cas2_mpd, max(Mortality$`fishing_pressure[Esp]`)))))

plot(c1_quant$fishing_pressures$year, c1_quant$fishing_pressures$Esp, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='U', main='Fishing pressure comparison: Esp')
lines(c1_sens1_quant$fishing_pressures$year, c1_sens1_quant$fishing_pressures$Esp, type='l', col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Mortality$year, cas2_mpd[[c]]$Mortality$`fishing_pressure[Esp]`, col=C2_color[c], lwd=1)
}

plot_legend()



max_val <- max(c1_quant$fishing_pressures$Wsp,
               c1_sens1_quant$fishing_pressures$Wsp,
               max(unlist(list.map(cas2_mpd, max(Mortality$`fishing_pressure[Wsp]`)))))

plot(c1_quant$fishing_pressures$year, c1_quant$fishing_pressures$Wsp, type='l', col='black', lwd=5, ylim=c(0, max_val), xlab='Year', ylab='U', main='Fishing pressure comparison: Wsp')
lines(c1_sens1_quant$fishing_pressures$year, c1_sens1_quant$fishing_pressures$Wsp, type='l', col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Mortality$year, cas2_mpd[[c]]$Mortality$`fishing_pressure[Wsp]`, col=C2_color[c], lwd=1)
}

plot_legend()
```


```{r plots_set_6, echo=FALSE}
# plot selectivity

ages <- seq(1, 17, 1)


max_val <- max(1,
               c1_quant$`Ogive parameter values`$`selectivity[Enspsl].all`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[Enspsl].all`)
plot(ages, c1_quant$`Ogive parameter values`$`selectivity[Enspsl].all`, type='l', col='black', lwd=5, ylim=c(0,max_val), xlab='Age', ylab='', main='Survey selectivity-at-age comparison: Enspsl')
lines(ages, c1_sens1_quant$`Ogive parameter values`$`selectivity[Enspsl].all`, col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(ages, cas2_mpd[[c]]$Enspsl$Values, col=C2_color[c], lwd=1)
}

plot_legend()



max_val <- max(1,
               c1_quant$`Ogive parameter values`$`selectivity[Wnspsl].all`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[Wnspsl].all`)
plot(ages, c1_quant$`Ogive parameter values`$`selectivity[Wnspsl].all`, type='l', col='black', lwd=5, ylim=c(0,max_val), xlab='Age', ylab='', main='Survey selectivity-at-age comparison: Wnspsl')
lines(ages, c1_sens1_quant$`Ogive parameter values`$`selectivity[Wnspsl].all`, col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(ages, cas2_mpd[[c]]$Wnspsl$Values, col=C2_color[c], lwd=1)
}

plot_legend()


max_val <- max(1,
               c1_quant$`Ogive parameter values`$`selectivity[Espsl].all`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[Espsl].all`)
plot(ages, c1_quant$`Ogive parameter values`$`selectivity[Espsl].all`, type='l', col='black', lwd=5, ylim=c(0,max_val), xlab='Age', ylab='', main='Survey selectivity-at-age comparison: Espsl')
lines(ages, c1_sens1_quant$`Ogive parameter values`$`selectivity[Espsl].all`, col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(ages, cas2_mpd[[c]]$Espsl$Values, col=C2_color[c], lwd=1)
}

plot_legend()


max_val <- max(1,
               c1_quant$`Ogive parameter values`$`selectivity[Wspsl].all`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[Wspsl].all`)
plot(ages, c1_quant$`Ogive parameter values`$`selectivity[Wspsl].all`, type='l', col='black', lwd=5, ylim=c(0,max_val), xlab='Age', ylab='', main='Survey selectivity-at-age comparison: Wspsl')
lines(ages, c1_sens1_quant$`Ogive parameter values`$`selectivity[Wspsl].all`, col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(ages, cas2_mpd[[c]]$Wspsl$Values, col=C2_color[c], lwd=1)
}

plot_legend()


max_val <- max(1,
               c1_quant$`Ogive parameter values`$`selectivity[CRsl].all`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[CRsl].all`)
plot(ages, c1_quant$`Ogive parameter values`$`selectivity[CRsl].all`, type='l', col='black', lwd=5, ylim=c(0,max_val), xlab='Age', ylab='', main='Survey selectivity-at-age comparison: CRsl')
lines(ages, c1_sens1_quant$`Ogive parameter values`$`selectivity[CRsl].all`, col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(ages, cas2_mpd[[c]]$CRsl$Values, col=C2_color[c], lwd=1)
}

plot_legend()


max_val <- max(1,
               c1_quant$`Ogive parameter values`$`selectivity[SAsl].all`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[SAsl].all`)
plot(ages, c1_quant$`Ogive parameter values`$`selectivity[SAsl].all`, type='l', col='black', lwd=5, ylim=c(0,max_val), xlab='Age', ylab='', main='Survey selectivity-at-age comparison: SAsl')
lines(ages, c1_sens1_quant$`Ogive parameter values`$`selectivity[SAsl].all`, col='grey', lwd=2.5)
for (c in 1:num_C2_models)
{
    lines(ages, cas2_mpd[[c]]$SAsl$Values, col=C2_color[c], lwd=1)
}

plot_legend()
```

