---
title: "Casal2 Test Case MPD comparisons for BCO"
output:
    html_document:
        toc: TRUE
        toc_depth: 2
        number_sections: TRUE

header-includes:
 - \usepackage{pdflscape}
 - \newcommand{\blandscape}{\begin{landscape}}
 - \newcommand{\elandscape}{\end{landscape}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# BCO comparison of CASAL and Casal2 model configurations

This document compares the results of at least 2 CASAL model configurations (base and at least one sensitivity) and up to 6 Casal2 model configurations (3 BetaDiff and 3 ADOL-C).

The CASAL model sensitivity 1 has a smaller minimisation tolerance value than the CASAL base model (1e-7 vs. 2e-5).

The Casal2 ADOL-C and BetaDiff low tolerance models have a smaller tolerance value than the CASAL base model (1e-7 vs. 2e-5).


## BCO model characteristics

The main characteristics of the Test Case BCO (ling) CASAL model are:

* one stock, ages 3 - 20
* one area
* years 1900 - 2019, projection year 2020
* one time step
* two stock categories: male and female
* natural mortality-at-age is 0.14 for males and 0.20 for females
* sex-specific logistic maturity
* sex-specific von Bertalanffy age-length relationship
* one length-weight relationship ($W = aL^b$)
* Beverton-Holt stock-recruitment relationship, with steepness (h) 0.75 and $\sigma_R$ 0.6
* no ageing error
* one survey with sex-specific age-based logistic selectivity
* three fisheries, all with sex-specific length-based logistic selectivity and retained/discarded catches

Observation data include:

* survey index
* fishery CPUE index
* fishery proportions-at-length

Parameters estimated include:

* B0
* male $L_{\infty}$
* male age-length CV2
* sex-specific survey logistic selectivity (a50 and ato95)
* YCS (recruitment deviations)


## R environment

```{r C1_C2_setup, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_set_up_R_env.R')
```


## CASAL and Casal2 model output

```{r CASAL, warning=FALSE}
source('../../R-functions/report_read_in_CASAL_MPD_files.R')
```


```{r Casal2}
source('../../R-functions/report_read_in_Casal2_MPD_files.R')
```



\newpage
\blandscape

## Tables

Tables of parameter estimates and objective function components for the CASAL and Casal2 model MPD results

```{r table_setup, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_set_up_tables.R')

options(scipen=999)
```


```{r huxtables_parameters, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_print_parameter_tables.R')

hux_C1_est_params
hux_C2_est_params
hux_C2_pd_est_params
```


```{r huxtables_obj_fun, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_print_obj_fun_tables.R')

hux_C1_obj_fun
hux_C2_obj_fun
```


```{r convergence_info, echo=FALSE}
source('../../R-functions/report_print_convergence_info.R')
```


```{r warnings_and_bounds, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_print_warnings.R')
```


```{r huxtables_close, echo=FALSE}
options(scipen=0)
```

\elandscape
\newpage



## Matching of outputs

Time series comparisons with CASAL base model results

```{r Time_Series_Match, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_time_series_match_function.R')


for (c in 1:num_C2_models)
{
    print(paste('Catch time series base model comparison for run', C2_subdir[c]))

    print(paste('Removals / Actual catches for line match:',
                time_series_match(c1_quant$removals$line, cas2_mpd[[c]]$Fishing$`actual_catch[FishingLine]`)))

    print(paste('Discarded catches for line match:',
                time_series_match(c1_quant$discards$line, cas2_mpd[[c]]$Fishing$`discards[FishingLine]`)))

    print(paste('Removals / Actual catches for pot match:',
                time_series_match(c1_quant$removals$pot, cas2_mpd[[c]]$Fishing$`actual_catch[FishingPot]`)))

    print(paste('Discarded catches for pot match:',
                time_series_match(c1_quant$discards$pot, cas2_mpd[[c]]$Fishing$`discards[FishingPot]`)))

    print(paste('Removals / Actual catches for rec match:',
                time_series_match(c1_quant$removals$rec, cas2_mpd[[c]]$Fishing$`actual_catch[Recreation]`)))

    print(paste('Discarded catches for rec match:',
                time_series_match(c1_quant$discards$rec, cas2_mpd[[c]]$Fishing$`discards[Recreation]`)))

    print('')
}
```

Derived quantities

SB0, SBcurrent, MSY, F_MSY, others...



## Plots

Comparison plots

```{r plots_base, echo=FALSE}

# c(bottom, left, top, right)
par(mar=c(4,4,2,1) + 0.1)

par(mfrow=c(1,2))

bl_lwd <- 6
gr_lwd <- 3
```


```{r plots_correlation, echo=FALSE}
source('../../R-functions/report_plot_correlation_matrices.R')
```


```{r plots_set_0, echo=FALSE}
source('../../R-functions/report_set_up_legend.R')


max_val <- max(c1_quant$SSBs$SSB,
               c1_sens1_quant$SSBs$SSB,
               max(unlist(list.map(cas2_mpd, max(SSB$SSB$values)))))

plot(c1_quant$SSBs$year, c1_quant$SSBs$SSB, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='Biomass (t)', main='SSB comparison')
lines(c1_sens1_quant$SSBs$year, c1_sens1_quant$SSBs$SSB, type='l', col='grey', lwd=gr_lwd)

for (c in 1:num_C2_models)
{
    lines(names(cas2_mpd[[c]]$SSB$SSB$values), cas2_mpd[[c]]$SSB$SSB$values, col=C2_color[c], lwd=1)
}

plot_legend()
```


```{r plots_set_1, echo=FALSE}

# Q:  how to get CASAL initial age structure?
# A:  use 'casal::extract.partition()''


ages <- seq(3, 20)

# omit the 'category' column
c2_mat <- array(0, dim=c(num_C2_models, dim(as.matrix(cas2_mpd[[1]]$Init$values[,-1]))))
for (c in 1:num_C2_models)
{
    c2_mat[c,,] <- as.matrix(cas2_mpd[[c]]$Init$values[,-1])
}

# plot initial numbers

max_val <- max(cas_part$initial,
               cas_sens1_part$initial,
               max(c2_mat))

plot(ages, cas_part$initial[1,], type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Age', ylab='', main='Initial numbers-at-age comparison (males and females)', sub='Dotted lines for female')
lines(ages, cas_part$initial[2,], col='black', lwd=bl_lwd, lty=3)

lines(ages, cas_sens1_part$initial[1,], col='grey', lwd=gr_lwd)
lines(ages, cas_sens1_part$initial[2,], col='grey', lwd=gr_lwd, lty=3)

for (c in 1:num_C2_models)
{
    lines(ages, c2_mat[c,1,], col=C2_color[c], lwd=1)
    lines(ages, c2_mat[c,2,], col=C2_color[c], lwd=1, lty=3)
}

plot_legend()



c2_mat <- array(0, dim=c(num_C2_models, dim(as.matrix(cas2_mpd[[1]]$state1$`2019`$values))))
for (c in 1:num_C2_models)
{
    c2_mat[c,,] <- as.matrix(cas2_mpd[[c]]$state1$`2019`$values)
}

# plot final numbers

max_val <- max(cas_part$final,
               cas_sens1_part$final,
               max(c2_mat))

plot(ages, cas_part$final[1,], type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Age', ylab='', main='Final numbers-at-age comparison (males and females)', sub='Dotted lines for female')
lines(ages, cas_part$final[2,], col='black', lwd=bl_lwd, lty=3)

lines(ages, cas_sens1_part$final[1,], col='grey', lwd=gr_lwd)
lines(ages, cas_sens1_part$final[2,], col='grey', lwd=gr_lwd, lty=3)

for (c in 1:num_C2_models)
{
    lines(ages, c2_mat[c,1,], col=C2_color[c], lwd=1)
    lines(ages, c2_mat[c,2,], col=C2_color[c], lwd=1, lty=3)
}

plot_legend()
```


```{r plots_set_2, echo=FALSE}
# plot YCS

max_val <- max(c1_quant$`Vector parameter values`$recruitment.YCS,
               c1_sens1_quant$`Vector parameter values`$recruitment.YCS,
               max(unlist(list.map(cas2_mpd, max(Recruitment$ycs_values)))))

plot(c1_quant$`Vector parameter values`$recruitment.YCS, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='', main='YCS comparison')
lines(c1_sens1_quant$`Vector parameter values`$recruitment.YCS, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Recruitment$ycs_values, col=C2_color[c], lwd=1)
}

plot_legend()



# plot true YCS

max_val <- max(c1_quant$true_YCS$true_YCS,
               c1_sens1_quant$true_YCS$true_YCS,
               max(unlist(list.map(cas2_mpd, max(Recruitment$true_ycs)))))

plot(c1_quant$true_YCS$year, c1_quant$true_YCS$true_YCS, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='', main='true YCS comparison')
lines(c1_sens1_quant$true_YCS$year, c1_sens1_quant$true_YCS$true_YCS, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Recruitment$ycs_years, cas2_mpd[[c]]$Recruitment$true_ycs, col=C2_color[c], lwd=1)
}

plot_legend()



# plot recruitment

max_val <- max(c1_quant$recruitments$recruitment,
               c1_sens1_quant$recruitments$recruitment,
               max(unlist(list.map(cas2_mpd, max(Recruitment$Recruits)))))

plot(c1_quant$recruitments$year, c1_quant$recruitments$recruitment, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='', main='Recruitment comparison')
lines(c1_sens1_quant$recruitments$year, c1_sens1_quant$recruitments$recruitment, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines((cas2_mpd[[c]]$Recruitment$ycs_years + cas2_mpd[[c]]$Recruitment$ssb_offset), cas2_mpd[[c]]$Recruitment$Recruits, col=C2_color[c], lwd=1)
}

plot_legend()
```


```{r plots_set_4, echo=FALSE}

# > names(c1_quant)
#  [1] "Scalar parameter values"        "Vector parameter values"
#  [3] "Ogive parameter values"         "Nuisance q's"
#  [5] "B0"                             "SSBs"
#  [7] "recruitments"                   "YCS"
#  [9] "true_YCS"                       "actual_catches"
# [11] "discards"                       "removals"
# [13] "fishing_pressures"              "potCPUE.fits"
# [15] "potSurveyAbd.fits"              "logbookLF.fits"
# [17] "potCommAF.fits"                 "potSurveyAF.fits"
# [19] "recCatchLF.fits"                "potCPUE.resids"
# [21] "potSurveyAbd.resids"            "logbookLF.resids"
# [23] "potCommAF.resids"               "potSurveyAF.resids"
# [25] "recCatchLF.resids"              "potCPUE.pearson_resids"
# [27] "potSurveyAbd.pearson_resids"    "logbookLF.pearson_resids"
# [29] "potCommAF.pearson_resids"       "potSurveyAF.pearson_resids"
# [31] "recCatchLF.pearson_resids"      "potCPUE.normalised_resids"
# [33] "potSurveyAbd.normalised_resids" "PotSurvey"
# [35] "selected_bio_Pot"


# plot removals (C1) / actual catch (C2)

max_val <- max(c1_quant$removals$line,
               c1_sens1_quant$removals$line,
               max(unlist(list.map(cas2_mpd, max(Fishing$`actual_catch[FishingLine]`)))))

plot(c1_quant$removals$year, c1_quant$removals$line, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='Catch (tonnes)', main='Removals / Actual catch comparison: line')
lines(c1_sens1_quant$removals$year, c1_sens1_quant$removals$line, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Fishing$year, cas2_mpd[[c]]$Fishing$`actual_catch[FishingLine]`, col=C2_color[c], lwd=1)
}

plot_legend()


# add percent difference




max_val <- max(c1_quant$removals$pot,
               c1_sens1_quant$removals$pot,
               max(unlist(list.map(cas2_mpd, max(Fishing$`actual_catch[FishingPot]`)))))

plot(c1_quant$removals$year, c1_quant$removals$pot, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='Catch (tonnes)', main='Removals / Actual catch comparison: pot')
lines(c1_sens1_quant$removals$year, c1_sens1_quant$removals$pot, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Fishing$year, cas2_mpd[[c]]$Fishing$`actual_catch[FishingPot]`, col=C2_color[c], lwd=1)
}

plot_legend()


max_val <- max(c1_quant$removals$rec,
               c1_sens1_quant$removals$rec,
               max(unlist(list.map(cas2_mpd, max(Fishing$`actual_catch[Recreation]`)))))

plot(c1_quant$removals$year, c1_quant$removals$rec, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='Catch (tonnes)', main='Removals / Actual catch comparison: recreational')
lines(c1_sens1_quant$removals$year, c1_sens1_quant$removals$rec, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Fishing$year, cas2_mpd[[c]]$Fishing$`actual_catch[Recreation]`, col=C2_color[c], lwd=1)
}

plot_legend()


# plot percent difference



# plot discarded catch

max_val <- max(c1_quant$discards$line,
               c1_sens1_quant$discards$line,
               max(unlist(list.map(cas2_mpd, max(Fishing$`discards[FishingLine]`)))))

plot(c1_quant$discards$year, c1_quant$discards$line, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='Catch (tonnes)', main='Discarded catch comparison: line')
lines(c1_sens1_quant$discards$year, c1_sens1_quant$discards$line, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Fishing$year, cas2_mpd[[c]]$Fishing$`discards[FishingLine]`, col=C2_color[c], lwd=1)
}

plot_legend()


max_val <- max(c1_quant$discards$pot,
               c1_sens1_quant$discards$pot,
               max(unlist(list.map(cas2_mpd, max(Fishing$`discards[FishingPot]`)))))

plot(c1_quant$discards$year, c1_quant$discards$pot, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='Catch (tonnes)', main='Discarded catch comparison: pot')
lines(c1_sens1_quant$discards$year, c1_sens1_quant$discards$pot, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Fishing$year, cas2_mpd[[c]]$Fishing$`discards[FishingPot]`, col=C2_color[c], lwd=1)
}

plot_legend()


max_val <- max(c1_quant$discards$rec,
               c1_sens1_quant$discards$rec,
               max(unlist(list.map(cas2_mpd, max(Fishing$`discards[Recreation]`)))))

plot(c1_quant$discards$year, c1_quant$discards$rec, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='Catch (tonnes)', main='Discarded catch comparison: recreational')
lines(c1_sens1_quant$discards$year, c1_sens1_quant$discards$rec, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Fishing$year, cas2_mpd[[c]]$Fishing$`discards[Recreation]`, col=C2_color[c], lwd=1)
}

plot_legend()


# plot percent difference


```


```{r plots_set_5, echo=FALSE}
# > names(cas2_mpd[[1]])
#  [1] "Init"                 "state1"               "estimated_values"
#  [4] "estimated_summary"    "obj_fun"              "Hess"
#  [7] "Covar"                "Corr"                 "Recruitment"
# [10] "Fishing"              "growth_length_at_age" "growth_weight_at_age"
# [13] "potSurveySel_male"    "potSurveySel_female"  "potSurvey"
# [16] "potCPUE"              "lgobookLF"            "potSurveyq"
# [19] "potCPUEq"             "SSB"                  "AFpotSurvey"
# [22] "AFpotFishing"         "rec.catch.LF"         "Ageing"
# [25] "MaturationSel_male"   "MaturationSel_female" "potFdis"
# [28] "recFdis"              "One"                  "messages_encountered"
# > names(cas2_mpd[[1]]$Fishing)
#  [1] "categories"                         "label"
#  [3] "m"                                  "relative_m_by_age"
#  [5] "time_step_proportions"              "sub_type"
#  [7] "year"                               "fishing_pressure[FishingLine]"
#  [9] "actual_catch[FishingLine]"          "retained_catch[FishingLine]"
# [11] "actual_retained_catch[FishingLine]" "discards[FishingLine]"
# [13] "discards_dead[FishingLine]"         "fishing_pressure[FishingPot]"
# [15] "actual_catch[FishingPot]"           "retained_catch[FishingPot]"
# [17] "actual_retained_catch[FishingPot]"  "discards[FishingPot]"
# [19] "discards_dead[FishingPot]"          "fishing_pressure[Recreation]"
# [21] "actual_catch[Recreation]"           "retained_catch[Recreation]"
# [23] "actual_retained_catch[Recreation]"  "discards[Recreation]"
# [25] "discards_dead[Recreation]"          "type"

# plot fishing pressures - line

max_val <- max(c1_quant$fishing_pressures$line,
               c1_sens1_quant$fishing_pressures$line,
               max(unlist(list.map(cas2_mpd, max(Fishing$`fishing_pressure[FishingLine]`)))))

plot(c1_quant$fishing_pressures$year, c1_quant$fishing_pressures$line, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='U', main='Fishing pressure comparison: line')
lines(c1_sens1_quant$fishing_pressures$year, c1_sens1_quant$fishing_pressures$line, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Fishing$year, cas2_mpd[[c]]$Fishing$`fishing_pressure[FishingLine]`, col=C2_color[c], lwd=1)
}

plot_legend()



# plot fishing pressures - pot

max_val <- max(c1_quant$fishing_pressures$pot,
               c1_sens1_quant$fishing_pressures$pot,
               max(unlist(list.map(cas2_mpd, max(Fishing$`fishing_pressure[FishingPot]`)))))

plot(c1_quant$fishing_pressures$year, c1_quant$fishing_pressures$pot, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='U', main='Fishing pressure comparison: pot')
lines(c1_sens1_quant$fishing_pressures$year, c1_sens1_quant$fishing_pressures$pot, type='l', col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Fishing$year, cas2_mpd[[c]]$Fishing$`fishing_pressure[FishingPot]`, col=C2_color[c], lwd=1)
}

plot_legend()



# plot fishing pressures - rec

max_val <- max(c1_quant$fishing_pressures$rec,
               c1_sens1_quant$fishing_pressures$rec,
               max(unlist(list.map(cas2_mpd, max(Fishing$`fishing_pressure[Recreation]`)))))

plot(c1_quant$fishing_pressures$year, c1_quant$fishing_pressures$rec, type='l', col='black', lwd=bl_lwd, ylim=c(0, max_val), xlab='Year', ylab='U', main='Fishing pressure comparison: recreational')
lines(c1_sens1_quant$fishing_pressures$year, c1_sens1_quant$fishing_pressures$rec, col='grey', lwd=gr_lwd)
for (c in 1:num_C2_models)
{
    lines(cas2_mpd[[c]]$Fishing$year, cas2_mpd[[c]]$Fishing$`fishing_pressure[Recreation]`, col=C2_color[c], lwd=1)
}

plot_legend()
```


```{r plots_set_6, echo=FALSE}

# plot survey selectivity

max_val <- max(1,
               c1_quant$`Ogive parameter values`$`selectivity[potSsel].male`, c1_quant$`Ogive parameter values`$`selectivity[potSsel].female`,
               c1_sens1_quant$`Ogive parameter values`$`selectivity[potSsel].male`, c1_sens1_quant$`Ogive parameter values`$`selectivity[potSsel].female`)

plot(ages, c1_quant$`Ogive parameter values`$`selectivity[potSsel].male`, type='l', col='black', lwd=bl_lwd, ylim=c(0,max_val), xlab='Age', ylab='', main='Pot survey selectivity-at-age comparison (males and females)', sub='Dotted lines for female')
lines(ages, c1_quant$`Ogive parameter values`$`selectivity[potSsel].female`, col='black', lwd=bl_lwd, lty=3)

lines(ages, c1_sens1_quant$`Ogive parameter values`$`selectivity[potSsel].male`, col='grey', lwd=gr_lwd)
lines(ages, c1_sens1_quant$`Ogive parameter values`$`selectivity[potSsel].female`, col='grey', lwd=gr_lwd, lty=3)

for (c in 1:num_C2_models)
{
    lines(ages, cas2_mpd[[c]]$potSurveySel_male$Values, col=C2_color[c], lwd=1)
    lines(ages, cas2_mpd[[c]]$potSurveySel_female$Values, col=C2_color[c], lwd=1, lty=3)
}

plot_legend()
```

