---
title: "Casal2 Test Case MCMC comparisons for BCO"
output:
    html_document:
        toc: TRUE
        toc_depth: 2
        number_sections: TRUE

header-includes:
 - \usepackage{pdflscape}
 - \newcommand{\blandscape}{\begin{landscape}}
 - \newcommand{\elandscape}{\end{landscape}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# BCO comparison of CASAL and Casal2 model configurations

This document compares the results of at least 2 CASAL model configurations (base and at least one sensitivity) and up to 6 Casal2 model configurations (3 BetaDiff and 3 ADOL-C).

The CASAL model sensitivity 1 has a smaller minimisation tolerance value than the CASAL base model (1e-7 vs. 2e-5).

The Casal2 ADOL-C and BetaDiff low tolerance models have a smaller tolerance value than the CASAL base model (1e-7 vs. 2e-5).


## BCO model characteristics

The main characteristics of the Test Case BCO (ling) CASAL model are:

* one stock, ages 3 - 20
* one area
* years 1900 - 2019, projection year 2020
* one time step
* two stock categories: male and female
* natural mortality-at-age is 0.14 for males and 0.20 for females
* sex-specific logistic maturity
* sex-specific von Bertalanffy age-length relationship
* one length-weight relationship ($W = aL^b$)
* Beverton-Holt stock-recruitment relationship, with steepness (h) 0.75 and $\sigma_R$ 0.6
* no ageing error
* one survey with sex-specific age-based logistic selectivity
* three fisheries, all with sex-specific length-based logistic selectivity and retained/discarded catches

Observation data include:

* survey index
* fishery CPUE index
* fishery proportions-at-length

Parameters estimated include:

* B0
* male $L_{\infty}$
* male age-length CV2
* sex-specific survey logistic selectivity (a50 and ato95)
* YCS (recruitment deviations)


## BCO model MCMC specifications

The CASAL MCMC options include

* start 0
* keep 2500
* length 4000000
* stepsize 0.02
* adaptive_stepsize True
* adapt_at 10000 50000 150000 250000 450000 750000 1400000
* df 2
* burn_in 1000
* proposal_t True

The Casal2 ADOL-C and BetaDiff MCMC options include

* start 0
* keep 2500
* length 4000000
* step_size 0.02
* adapt_stepsize_at 10000 50000 150000 250000 450000 750000 1400000
* proposal_distribution t
* adapt_stepsize_method double_half


## R environment

```{r C1_C2_setup, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_set_up_R_env.R')
```


## CASAL and Casal2 model output

```{r CASAL, warning=FALSE}
# source('../../R-functions/report_read_in_CASAL_MPD_files.R')
source('../../R-functions/report_read_in_CASAL_MCMC_files.R')
```


```{r Casal2}
source('../../R-functions/report_read_in_Casal2_MPD_files.R')
source('../../R-functions/report_read_in_Casal2_MCMC_files.R')
```


## MCMC diagnostics

### Diagnostics for the CASAL and Casal2 model MCMC results

For the diagnostics below, the last 1000 samples for each chain are input into the diagnostic functions.


```{r CASAL_MCMC_diag, echo=FALSE, message=FALSE}
plot_file_label <- 'BCO'

num_samples_to_use <- 1000
num_to_subsample   <- 1
MCMC_params_rename <- c('objective_function', 'SB0', 'Male_Linf', 'SelMale_a50', 'SelFemale_a50', 'YCS_last')

C1_MCMC_params <- c('neg_ln_posterior', 'initialization.B0', 'size_at_age.Linf_male.1', 'selectivity[potSsel].male.1', 'selectivity[potSsel].female.1', 'recruitment.YCS.118')

source('../../R-functions/report_CASAL_MCMC_diagnostics.R')
```


```{r Casal2_MCMC_diag, echo=FALSE, message=FALSE}
C2_MCMC_params <- c('objective_score', 'process[Recruitment].b0', 'age_length[asMm0].linf', 'selectivity[potSurveySel_male].a50', 'selectivity[potSurveySel_female].a50', 'process[Recruitment].ycs_values{2014}')

source('../../R-functions/report_Casal2_MCMC_diagnostics.R')
```


```{r plots_base, echo=FALSE}
# c(bottom, left, top, right)
par(mar=c(4,4,2,1) + 0.1)

par(mfrow=c(1,1))
```


```{r plot_MCMC_coda_diag, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_plot_MCMC_coda_diagnostics.R')
```


```{r plot_MCMC_ggmcmc_diag, echo=FALSE, warning=FALSE, message=FALSE}
source('../../R-functions/report_plot_MCMC_ggmcmc_diagnostics.R')
```

