\defComLab{Likelihood}{Define an object of type \emph{Likelihood}}.
\defRef{sec:Likelihood}
\label{syntax:Likelihood}

\subsubsection{Likelihood of type Binomial}
\commandlabsubarg{Likelihood}{type}{Binomial}.
\label{syntax:Likelihood-Binomial}

The Binomial type has no additional subcommands.
\subsubsection{Likelihood of type Binomial\_Approx}
\commandlabsubarg{Likelihood}{type}{Binomial\_Approx}.
\label{syntax:Likelihood-BinomialApprox}

The Binomial\_Approx type has no additional subcommands.
\subsubsection{Likelihood of type Dirichlet}
\commandlabsubarg{Likelihood}{type}{Dirichlet}.
\label{syntax:Likelihood-Dirichlet}

The Dirichlet type has no additional subcommands.
\subsubsection{Likelihood of type Log\_Normal}
\commandlabsubarg{Likelihood}{type}{Log\_Normal}.
\label{syntax:Likelihood-LogNormal}

The Log\_Normal type has no additional subcommands.
\subsubsection{Likelihood of type Log\_Normal\_With\_Q}
\commandlabsubarg{Likelihood}{type}{Log\_Normal\_With\_Q}.
\label{syntax:Likelihood-LogNormalWithQ}

The Log\_Normal\_With\_Q type has no additional subcommands.
\subsubsection{Likelihood of type Multinomial}
\commandlabsubarg{Likelihood}{type}{Multinomial}.
\label{syntax:Likelihood-Multinomial}

The Multinomial type has no additional subcommands.
\subsubsection{Likelihood of type Normal}
\commandlabsubarg{Likelihood}{type}{Normal}.
\label{syntax:Likelihood-Normal}

The Normal type has no additional subcommands.
\subsubsection{Likelihood of type Pseudo}
\commandlabsubarg{Likelihood}{type}{Pseudo}.
\label{syntax:Likelihood-Pseudo}

The Pseudo type has no additional subcommands.
