\defComLab{Estimate\_Transformation}{Define an object of type \emph{Estimate\_Transformation}}.
\defRef{sec:EstimateTransformation}
\label{syntax:EstimateTransformation}

\defSub{label}{Label for the transformation block}
\defType{String}
\defDefault{No default}

\defSub{type}{The type of transformation}
\defType{String}
\defDefault{No default}

\defSub{transform\_with\_jacobian}{Apply Jacobian during transformation}
\defType{Boolean}
\defDefault{true}

\subsubsection{Estimate\_Transformation of type Average\_Difference}
\commandlabsubarg{Estimate\_Transformation}{type}{Average\_Difference}.
\defRef{sec:EstimateTransformation-AverageDifference}
\label{syntax:EstimateTransformation-AverageDifference}

\defSub{theta2}{The label of the @estimate block relating to the $\theta\_2$ parameter in the transformation. See the User Manual for more information}
\defType{String}
\defDefault{No default}

\defSub{theta1}{The label of @estimate block relating to the $\theta\_1$ parameter in the transformation. See the User Manual for more information}
\defType{String}
\defDefault{No default}

\subsubsection{Estimate\_Transformation of type Inverse}
\commandlabsubarg{Estimate\_Transformation}{type}{Inverse}.
\defRef{sec:EstimateTransformation-Inverse}
\label{syntax:EstimateTransformation-Inverse}

\defSub{estimate\_label}{The label of estimate block to apply transformation. Defined as $\theta\_1$ in the documentation}
\defType{String}
\defDefault{No default}

\subsubsection{Estimate\_Transformation of type Log}
\commandlabsubarg{Estimate\_Transformation}{type}{Log}.
\defRef{sec:EstimateTransformation-Log}
\label{syntax:EstimateTransformation-Log}

\defSub{estimate\_label}{Label of estimate block to apply transformation. Defined as $\theta\_1$ in the documentation}
\defType{String}
\defDefault{No default}

\subsubsection{Estimate\_Transformation of type Log\_Sum}
\commandlabsubarg{Estimate\_Transformation}{type}{Log\_Sum}.
\defRef{sec:EstimateTransformation-LogSum}
\label{syntax:EstimateTransformation-LogSum}

\defSub{theta2}{The label of the @estimate block relating to the $\theta\_2$ parameter in the transformation. See the User Manual for more information}
\defType{String}
\defDefault{No default}

\defSub{theta1}{The label of @estimate block relating to the $\theta\_1$ parameter in the transformation. See the User Manual for more information}
\defType{String}
\defDefault{No default}

\subsubsection{Estimate\_Transformation of type Orthogonal}
\commandlabsubarg{Estimate\_Transformation}{type}{Orthogonal}.
\defRef{sec:EstimateTransformation-Orthogonal}
\label{syntax:EstimateTransformation-Orthogonal}

\defSub{theta2}{The label of the @estimate block relating to the $\theta\_2$ parameter in the transformation. See the User Manual for more information}
\defType{String}
\defDefault{No default}

\defSub{theta1}{The label of @estimate block relating to the $\theta\_1$ parameter in the transformation. See the User Manual for more information}
\defType{String}
\defDefault{No default}

\subsubsection{Estimate\_Transformation of type Simplex}
\commandlabsubarg{Estimate\_Transformation}{type}{Simplex}.
\defRef{sec:EstimateTransformation-Simplex}
\label{syntax:EstimateTransformation-Simplex}

\defSub{lower\_bound}{The empirical lower bound for the simplex transformed. This should be -Inf but some of the minimisers do not allow that}
\defType{Real number}
\defDefault{true}

\defSub{upper\_bound}{The empirical upper bound for the simplex transformed. This should be Inf but some of the minimisers do not allow that}
\defType{Real number}
\defDefault{true}

\subsubsection{Estimate\_Transformation of type Square\_Root}
\commandlabsubarg{Estimate\_Transformation}{type}{Square\_Root}.
\defRef{sec:EstimateTransformation-SquareRoot}
\label{syntax:EstimateTransformation-SquareRoot}

\defSub{estimate\_label}{The label of the estimate block to apply transformation. Defined as $\theta\_1$ in the documentation}
\defType{String}
\defDefault{No default}

\subsubsection{Estimate\_Transformation of type Sum\_To\_One}
\commandlabsubarg{Estimate\_Transformation}{type}{Sum\_To\_One}.
\defRef{sec:EstimateTransformation-SumToOne}
\label{syntax:EstimateTransformation-SumToOne}

\defSub{estimate\_labels}{The label for the estimates for the sum to one transformation}
\defType{Vector of strings}
\defDefault{No default}

\defSub{upper\_bound}{The empirical upper bounds for the transformed parameters. There should be one less bound than parameters}
\defType{Vector of real numbers (estimable)}
\defDefault{true}

\defSub{lower\_bound}{The empirical lower bound for the transformed parameters. There should be one less bound than parameters}
\defType{Vector of real numbers (estimable)}
\defDefault{true}

