\section{Introduction\label{sec:Introduction}}

The Contributors Guide provides an overview for those who wish to contribute to the \CNAME\ source code and compile the latest version of \CNAME. \CNAME\ is open source and the Development Team encourages users to submit and contribute changes, bug fixes, or enhancements. This document is intended to provide a guide for those who wish to undertake the these tasks. 

The \CNAME\ source code is hosted on GitHub, and can be found at \url{https://github.com/NIWAFisheriesModelling/CASAL2}.

A release bundle which includes a binary (both windows and Linux operating systems), manual, examples, \R\  package and other help guides can be downloaded at \url{https://github.com/NIWAFisheriesModelling/CASAL2/releases}. 

\CNAME\ release versions on the front page of the repository can lag behind the latest source code because uploading these is manual process. It is possible to get an executable for both windows and Linux (Ubuntu 20) using the GitHub actions artefacts \href{https://github.com/NIWAFisheriesModelling/CASAL2/actions}{see here}. To obtain these, log into your GitHub account then click on the latest successfully checked commit (indicated by a green tick), scroll to the bottom and download either \texttt{Casal2-Linux-build} or \texttt{Casal2-Windows-build}. These files are only saved for 30 days so if there hasn't been a commit in a while, they may not be available.

To help maintain the quality of the code base, the development team has created this guide to assist contributors in understanding how to access the source code, how to compile and build, and the guidelines for submitting code changes to the \CNAME\ Development Team.  

This guide covers basics such as setting up a GitHub profile, to using GitHub, and all the way to compiling and modifying source code. 

As a suggestion to contributors, it may be helpful to contact the Development Team before you begin making the change. They can assist with ideas on how best to add functionality. Either contact them directly or open an issue at \url{https://github.com/NIWAFisheriesModelling/CASAL2/issues}. 

If you have any questions or require more information, please contact the \CNAME\ Development Team at \email.